---
slug: "news-august-2026"
title: rOpenSci News Digest, August 2026
author:
  - The rOpenSci Team
date: '2026-08-28'
tags:
  - newsletter
description: New editors; champions program; Sovereign Tech Fellowship; Quinceañera; new packages and package news
params:
  last_newsletter: '2026-06-30'
  doi: "10.59350/4chwm-mtm35"
---

```{r setup, include=FALSE}
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
  "/blog/%s/%s/%s/%s",
  lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
  stringr::str_pad(
    lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    2,
    "0",
    side = "left"
  ),
  stringr::str_pad(
    lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    2,
    "0",
    side = "left"
  ),
  rmarkdown::yaml_front_matter(knitr::current_input())$slug
)
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(trimws(...), sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ


### Champions Program update

### R-Universe updates

:muscle: R-universe has started building and checking packages for Windows ARM64. 
Read more in our [tech note](/blog/2026/08/06/r-universe-winarm/).

:dna: R-universe is now part of the infrastructure of the [Bioconductor submission process](https://blog.bioconductor.org/posts/2026-06-15-new-submission-process-with-Runiverse/).

## We're still celebrating our 15th anniversary! 🎉

In July, we started to share stories from members of our community about their experiences with rOpenSci. Our first story features [Eunseop Kim](/author/eunseop-kim/) and his connection with rOpenSci.
Read it on our blog: [From Peer Review to Mentorship: My rOpenSci Story](/blog/2026/07/14/15yo-eunseop-kim/)
Stay tuned for more stories from our community as we continue celebrating 15 years of rOpenSci!

### Coworking 

Read [all about coworking](/blog/2023/06/21/coworking/)!

- Tuesday September 1st, 14:00 Europe Central (12:00 UTC) ["Getting to Know SORTEE"](/events/coworking-2026-09/), with [Steffi LaZerte](/author/steffi-lazerte) and cohost [Ed Ivimey-Cook](/author/ed-ivimey-cook/).
  - Visit [SORTEE](https://sortee.org/) (Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology).
  - Meet co-host, Ed Ivimey-Cook, and learn more about SORTEE and how you might get involved.
 
And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!


## Software :package:


```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$Package,
    description = cran_unquote(entry$Title),
    details = cran_unquote(entry$Description),
    on_cran = !is.null(entry$cran_url),
    on_bioc = !is.null(entry$`_bioc`),
    onboarding = entry$`_metadata`$review$url %||% "",
    url = entry$RemoteUrl,
    maintainer = sub(" >.*", "", entry$Maintainer)
  )
}

registry <- "https://ropensci.r-universe.dev/api/packages/" |>
  jsonlite::read_json() |>
  purrr::map_df(tidy_package)

since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 21
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" |>
    sprintf(commits[[i]]$sha) |>
    jsonlite::read_json() |>
    purrr::map_df(function(x) {
      tibble::tibble(package = x$package, url = x$url, branch = x$branch)
    })
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !package %in% "monarch",
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE, eval=TRUE}
new$onboarding[
  new$package == "hdcuremodels"
] <- "https://github.com/ropensci/software-review/issues/692"
packages <- split(new, seq(nrow(new)))
packages <- rev(packages)
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) |>
    purrr::pluck("download_url")

  if (package$package == "eDNAjoint") {
    desc_link <- "https://raw.githubusercontent.com/ropensci/eDNAjoint/26fb2d22dcd4314cdf7d170812c6247ae51f25a0/DESCRIPTION"
  }

  withr::with_tempfile("tf", {
    download.file(desc_link, tf)
    desc <<- desc::desc(file = tf)
  })
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {
    any(x$role %in% "aut") && all(x$role != "cre")
  })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {
    any(x$role %in% "rev") && all(x$role != "cre")
  })
  rev <- purrr::map_chr(rev, function(x) {
    paste(c(x$given, x$family), collapse = " ")
  })
  if (package$package == "pvEBayes") {
    rev <- c("Kathryn Doering", "Collin Cademartori")
  }
  if (package$package == "nycOpenData") {
    rev <- c("Haolin Dong", "Michael Pascale")
  }
  maintainer <- purrr::keep(authors, function(x) {
    any(x$role %in% "cre")
  })
  maintainer <- paste(
    c(maintainer[[1]]$given, maintainer[[1]]$family),
    collapse = " "
  )
  author_string <- sprintf("developed by %s", maintainer)

  if (length(aut) > 0) {
    author_string <- paste0(
      author_string,
      sprintf(" together with %s", nice_string(aut))
    )
  }

  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package,
    package$package,
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )

  string <- gsub("doi\\:", "https://doi.org/", string)

  if (package$on_cran) {
    string <- paste0(
      string,
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string,
      sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(
      string,
      sprintf("It has been [reviewed](%s)", package$onboarding)
    )
    if (package$package == "fireexposuR") {
      rev <- c("Sherry Zhang ", "Ronny A. Hernández Mora")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  string <- sub("twitter", "social media", string)
  paste("+", string)
}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry |>
  dplyr::rowwise() |>
  dplyr::mutate(
    owner = strsplit(urltools::path(url), "/")[[1]][1],
    repo = strsplit(urltools::path(url), "/")[[1]][2]
  ) |>
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if (
    length(info) == 0 ||
      anytime::anytime(info[[1]]$published_at) < last_newsletter
  ) {
    return(NULL)
  }

  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

The writexl package has a [new maintainer](https://github.com/ropensci/writexl/pull/98#issuecomment-4191858158), Bill Denney.
NLMR is now maintained by [Jakub Nowosad](https://github.com/ropensci/NLMR/issues/116#issuecomment-4280937012).

## Software Peer Review

```{r pkgs}
#| cache: true
#| results: 'asis'
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub(
    "Package:",
    "",
    lines[grepl("^Package", lines)][1]
  ))
  description <- stringr::str_squish(sub(
    "Title:",
    "",
    lines[grepl("^Title", lines)][1]
  ))
  if (name == "read.abares") {
    description <- "Provides simple downloading, parsing and importing of Australian
 Bureau of Agricultural and Resource Economics and Sciences (ABARES) data sources"
  }
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login

  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!startsWith(url, "https")) {
    url <- paste0("https://", url)
  }
  sprintf("[%s](%s)", name, url)
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else(
      "stats" %in% purrr::map_chr(issue$labels, "name"),
      " (Stats).",
      ""
    )
  )

  dplyr::rowwise(df) |>
    dplyr::mutate(
      text = sprintf(
        "    * [%s](%s), %s. Submitted by %s. %s",
        name,
        url,
        description,
        user,
        stats
      )
    )
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state,
    .limit = Inf,
    .progress = FALSE
  )

  issues <- purrr::keep(issues, \(x) {
    label %in% purrr::map_chr(x[["labels"]], "name")
  })

  purrr::map_df(issues, tidy_issue)
}

active_issues <- purrr::map_df(
  c(
    "1/editor-checks",
    "2/seeking-reviewer(s)",
    "3/reviewer(s)-assigned",
    "4/review(s)-in-awaiting-changes",
    "5/awaiting-reviewer(s)-response",
    "6/approved"
  ),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label = "6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  }

  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) |>
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf(
  "There are %s recently closed and active submissions",
  english(nrow(issues))
)
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(
    issues$label == label,
    na.rm = TRUE
  )))
  url <- glue::glue(
    'https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A%22{label}%22'
  )
  sprintf(
    "* %s at ['%s'](%s):\n\n %s",
    no,
    label,
    url,
    paste0(
      issues$text[!is.na(issues$label)][issues$label == label],
      collapse = "\n\n"
    )
  )
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(
        sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE),
        count_label
      ),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->
```{r blog, cache=TRUE}
#| results: 'asis'
parse_one_post <- function(path) {
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))

  if (lines[2] == "headless: true") {
    return(NULL)
  }

  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)

  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in%
      yaml$tags &&
      !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags &&
      !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
  )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[
  !paths %in%
    c(
      "../_index.md",
      "../2021-02-03-targets/raw_data_source.md",
      "../2021-02-03-targets/README.md"
    )
]
paths <- grep("index", paths, value = TRUE)
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }

  post <- dir[which(dir[["language"]] == main_language), ]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(
      string,
      ". ",
      sub("[\\.\\!\\?]$", "", post$description),
      "."
    )
    if (url == "/blog/2025/09/18/markdown-programmatic-parsing/") {
      string <- paste(string, ".")
    }
  } else {
    string <- paste0(string, ".")
  }
  if (post$title == "How to Assess Usage of your Package") {
    string <- paste0(
      string,
      " This post was featured in the [R Weekly Highlights podcast](https://serve.podhome.fm/episodepage/r-weekly-highlights/216) hosted by Eric Nantz and Mike Thomas."
    )
  }

  other_langs <- dir[which(dir[["language"]] != main_language), ]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf(
        "<a href='%s' lang='%s'>%s (%s)</a>",
        .x[["url"]],
        .x[["language"]],
        .x[["title"]],
        .x[["language"]]
      )
    ) |>
      toString()
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  if (post$socialImg != "") {
    if (!startsWith(post$socialImg, "https")) {
      img_file <- fs::path_file(post$socialImg)
      download.file(
        sprintf("https://ropensci.org/%s", post$socialImg),
        img_file
      )
      img_file |>
        magick::image_read() |>
        magick::image_scale("400x") |>
        magick::image_write(img_file)
    } else {
      img_file <- post$socialImg
    }

    string <- paste0(
      string,
      "\n```{=html}\n",
      sprintf(
        '{{< figure class="center" src="%s" alt="%s" width="400" >}}\n\n',
        img_file,
        post$socialAlt
      ),
      "\n```\n"
    )
  }

  string
}

software_review <- posts[purrr::map_lgl(
  posts,
  ~ any(.x[["software_peer_review"]])
)]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~ any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~ any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

- [charlatan](https://docs.ropensci.org/charlatan), create fake data in R.
  [Issue for volunteering](https://github.com/ropensci/charlatan/issues/150).

### Calls for contributions

Refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.

## Package development corner

Some useful information for R package developers.
:eyes:

## From Rd files to Quarto

Edgar Ruiz from Posit released [pkgsite](https://opensource.posit.co/blog/2026-06-18_pkgsite-0-1-0/),
a package for converting your package's `.Rd` files to Quarto.
It creates qmd files that you can integrate as you want in a Quarto website.

## Mutation testing, fuzzy testing

First of all, a reminder in case you confuse the two concepts...

👽️ Mutation testing: you run tests on mutated version of the *code*. 

🚮 Fuzzy testing: you run tests with all sorts of *inputs* to the code.

At useR! 2026, both topics were covered:

- mutation testing in [Beyond Code Coverage: Mutation Testing in R with mutator](https://docs.google.com/presentation/d/1T5iK0gRBFF869Q6eirtmYVV-5ABD4EN-WBJI5gl5EUY/edit?slide=id.g3efebd0f55a_0_99#slide=id.g3efebd0f55a_0_99) by Assanali Amandykov, Pierre Donat-Bouillud.
- fuzzy testing in [Fuzz-testing R-based research software for robustness](https://events.digital-research.academy/event/109/contributions/430/attachments/187/407/slides-v2.pdf) by Marco Colombo.

## roxyreqs

Also at useR!, Moritz Lang and collaborators introduced [roxyreqs](https://github.com/mnlang/roxyreqs), a package for adding roxygen2-like documentation to testthat.
[Slides](https://events.digital-research.academy/event/109/contributions/471/attachments/69/201/roxyreqs-user2026.pdf).

## checktor, a new helper for CRAN submissions

If you want to submit your package to CRAN, you can get help through the [CRAN cookbook](https://contributor.r-project.org/cran-cookbook/), the [collaborative list maintained by ThinkR](https://github.com/ThinkR-open/prepare-for-cran) and now a new package, checktor by James Balamuta!
Read more in [James' post](https://blog.thecoatlessprofessor.com/programming/r/the-check-passed-the-reviewer-didnt/).

## Interesting AI reads

- [Recommendations When Using LLM-backed Generative AI Systems for FOSS Contributions](https://sfconservancy.org/llm-gen-ai/llm-backed-generative-ai-recommendations.html) by Software Freedom Conservancy, shared by Will Gearty.
- [Who cleans up after the vibe-coding party?](https://www.ft.com/content/cec8df9e-b43b-4cd1-8feb-c07e804e8d33) by Sam Learner in the Financial Times.
- [AI Tooling and Open Source](https://niccrane.com/posts/ai-tooling-open-source/) in which Nic Crane discusses "how AI tooling affects open source, the actions maintainers have been taking to address the less positive aspects, and emerging policies that open source projects are implementing around the topic of AI-generated pull requests".

## Last words

Thanks for reading!
If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org).
This guide will help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways such as through sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so though our signup form](/news/).
Until it's time for our next newsletter, you can keep in touch with us through our [website](/), [Mastodon](https://hachyderm.io/@rOpenSci), or [LinkedIn](https://www.linkedin.com/company/ropensci/).
See you soon!
