---
slug: "news-july-2024"
title: rOpenSci News Digest, July 2024
author:
  - The rOpenSci Team
date: '2024-07-19'
tags:
  - newsletter
description: keywords from the content
params:
  last_newsletter: "2024-06-21"
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### Announcing New Software Peer Review Editors: Beatriz Milz and Margaret Siple

We are excited to welcome Beatriz Milz and Margaret Siple to our team of [Associate Editors](https://ropensci.org/software-review/#editors) for rOpenSci [Software Peer Review](https://ropensci.org/software-review/). They join Laura DeCicco, Julia Gustavsen, Anna Krystalli, Mauro Lepore, Noam Ross, Maëlle Salmon, Emily Riederer, Adam Sparks, and Jeff Hollister. 

Meet Beatriz and Margaret in their [introduction blog post](/blog/2024/07/03/editors2024/).
Welcome on board to both of them, thank you for your service!

### A fresh new look for R-universe!

You might have noticed that R-universe got a big refresh. :icecream:
Read all about this [big overhaul of the interface](/blog/2024/06/12/runiverse-frontend/).

### Resources from the rOpenSci community at useR! 2024

While some video recordings have not been posted on the [useR! YouTube channel](https://www.youtube.com/@useRConference_global) yet, some slidedecks and materials are already available.

- [Maëlle Salmon](/author/maëlle-salmon/)'s keynote talk ["How your code might get rusty"](https://sched.co/1c9Jy). [Slides](https://user-maelle.netlify.app/), [recording](https://www.youtube.com/watch?v=ldko1TPC1aA).
- [Jeroen Ooms](/author/jeroen-ooms/)' talk  ["Navigating the R Ecosystem Using R-Universe"](https://sched.co/1c90h). [Slides](https://jeroen.github.io/user2024/#/title-slide).
- [Jon Harmon](/author/jon-harmon/)'s virtual talk ["Learning Together at the Data Science Learning Community"](https://www.youtube.com/watch?v=8jH5wfuDx-g).
- [Paola Corrales](/author/pao-corrales/)' and [Elio Campitelli](/author/elio-campitelli/)'s tutorial ["Efficient Data Analysis with data.table"](https://sched.co/1c8yF). [Website](https://paocorrales.github.io/intro-datatable/).
- [Elio Campitelli](/author/elio-campitelli)'s talk ["Building Bilingual Bridges with Multilingual Manuals"](https://sched.co/1c8yj). [Slides](https://static.sched.com/hosted_files/userconf2024/3d/UseR2024-1.pdf).
- [Hugo Gruson](/author/hugo-gruson)'s talk ["
Building Interoperability in Existing Software Ecosystems with S3 Classes"](https://sched.co/1c8z3). [Slides](https://static.sched.com/hosted_files/userconf2024/ff/2024-07-09_Gruson%20-%20Building%20Interoperability%20with%20S3%20Classes.pdf). Also his poster ["A reproducible analysis of CRAN Task Views to understand the state of an R package ecosystem"](https://static.sched.com/hosted_files/userconf2024/ef/HugoGruson_useR_poster.pdf).
- [Lluís Revilla](/author/lluís-revilla-sancho/)'s and Henrik Bengtsson's [poster](https://sched.co/1cDsG) about CRAN packages archived and a the [cranhaven.org](http://cranhaven.org/) R-universe created to reduce the impact of that on users.
- [Will Landau](/author/will-landau)'s and Charlie Gao's talk ["Moju-Kapu: How {Mirai} and {Crew} Are Powering the Next Generation of Parallel Computing in R"](https://sched.co/1c8ya). [Slides](https://static.sched.com/hosted_files/userconf2024/8a/moju-kapu.html).
- [Binod Jung Bogati](/blog/2024/02/15/champions-program-champions-2024/#binod-jung-bogati)'s talks ["Generate Raw Synthetic Dataset for Clinical Trial - Binod Jung Bogati, Numeric Mind"](https://sched.co/1c8w7), [slides](https://static.sched.com/hosted_files/userconf2024/ff/Generate%20Raw%20Synthetic%20Data%20for%20Clinical%20Trial%20-%20Binod%20Jung%20Bogati%20-%20useR2024.pdf); and ["Translate R for Global Reach"](https://sched.co/1c90Y).

### Coworking

Read [all about coworking](/blog/2023/06/21/coworking/)!

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.

- [Tuesday, August 6th, 9:00 America Pacific (16:00 UTC)](/events/coworking-2024-08), Building your first R package with cohost [Carolina Pradier](/author/carolina-pradier/) and [Steffi LaZerte](/author/steffi-lazerte).
    - Explore how to make R packages
    - Plan out that package you’ve always wanted to create
    - Chat with our cohost about tips and tricks for making your first R package

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE,eval=FALSE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$socialImg != "") {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt)
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  if (length(others) != length(posts)) cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Use cases

```{r usecases}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
last_date <- anytime::anydate("2024-06-01")
usecases <- dplyr::filter(usecases, date >= last_date)
usecases <- split(usecases, seq(nrow(usecases)))
```

`r snakecase::to_sentence_case(english(length(usecases)))` use case`r if (length(usecases) > 1) "s"` of our packages and resources ha`r if (length(usecases) > 1) "ve" else "s"` been reported recently.

```{r usecases2, results='asis'}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", sub("\\.$", "", usecase$title), usecase$url, usecase$reporter)
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)!


## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

* [historydata](https://docs.ropensci.org/historydata/), datasets for historians. [Issue for volunteering](https://github.com/ropensci/historydata/issues/23).

* [USAboundaries](https://docs.ropensci.org/USAboundaries/) (and USAboundariesdata), historical and contemporary boundaries of the United States of America . [Issue for volunteering](https://github.com/ropensci/USAboundaries/issues/50).

### Calls for contributions

* [Help make qualtRics better! Code contributions wanted](/blog/2024/03/01/qualtrics-call-help/)

* [Help make assertr better! Come close issues](/blog/2024/02/27/assertr-call-help/)

* [Help users and developers of targets by answering questions!](/blog/2024/02/29/targets-call-help/)

* [Help make waywiser better! User requests wanted](/blog/2024/02/19/waywiser-call-help/)

Also refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.

## Package development corner

Some useful tips for R package developers. :eyes:

### Last call: your opinion on the CRAN submission process!

Shared by Lluís Revilla and Heather Turner in our Slack workspace, a crucial survey ending today!

>  "If you have R package development experience and would like to share your thoughts on the CRAN submission process, please fill this short survey from the CRAN Cookbook project!"

Please find the [Google form](https://forms.gle/umdew9KHWeQSehq8A) and read more about the exciting cookbook project in this [post by Jasmine D.](https://www.linkedin.com/pulse/improving-skills-r-package-maintainers-support-from-consortium-daly-tlt4e).

### Robust type-checking with r-lib

Don't miss this insightful short post by Josiah Parry, ["Type safe(r) R code"](https://josiahparry.com/posts/2024-06-30-type-safety/).
A related older blog post is ["Checking the inputs of your R functions"](https://blog.r-hub.io/2022/03/10/input-checking/) by Hugo Gruson, Sam Abbott, Carl Pearson.

### The one with all the useR! links

The useR! 2024 conference featured quite a few talks relevant to package development, beside the talks we mentioned in the [HQ section](#resources-from-the-ropensci-community-at-user-2024).
Not all recordings are available yet, but make sure to check out the useR! YouTube channel.

If we missed any relevant content, please [get in touch](https://github.com/ropensci-org/monthly/issues) so we might add missing pieces to our next newsletter!

### Retrospectives

Kurt Hornik gave a keynote talks on [More than 25 years of CRAN](https://userconf2024.sched.com/event/1c9Jm/keynote-more-than-25-years-of-cran-kurt-hornik-wirtschaftsuniversitat-wien?iframe=no&w=100%&sidebar=yes&bg=no) ([slides](https://static.sched.com/hosted_files/userconf2024/41/user2024_CRAN.pdf)) and Torsten Hothorn on [Some things you can’t read from
a NEWS file](https://userconf2024.sched.com/event/1c9JC/keynote-some-things-you-cant-read-from-a-news-file-torsten-hothorn-university-of-zurich?iframe=yes&w=100%&sidebar=yes&bg=no) ([slides](Some things you can’t read from
a NEWS file)) about maintaining a package for decades.

#### On validation of R packages

Coline Zeballos and Yann Féhat from the R Validation Hub discussed [how to support (pharma) companies with validation of R Packages](https://userconf2024.sched.com/event/1dyXW/supporting-companies-with-validation-of-r-packages-yann-feat-mainanalytics-gmbh-coline-zeballos-r-validation-hub?iframe=no&w=100%&sidebar=yes&bg=no). [Slides](https://static.sched.com/hosted_files/userconf2024/4a/Supporting%20Companies%20With%20Validation%20of%20R%20Packages_%20A%20Regulatory%20Repository.pdf). They use a toolset based on [r-hub/repos](https://github.com/r-hub/repos) and the [riskmetric](https://pharmar.github.io/riskmetric/) package.

Szymon Maksymiuk and Lorenzo Braschi presented a [Deep Dive Into Industry R
Package Quality Assessment](https://userconf2024.sched.com/event/1c8z7/deep-dive-into-industry-r-package-quality-assessment-szymon-maksymiuk-lorenzo-braschi-roche?iframe=no). Beside introducing the concepts, they mentioned three open-source R packages that they created: [checked](https://github.com/Genentech/checked) for running reverse dependencies checks; [covtracer](https://genentech.github.io/covtracer/) for contextualizing tests, built using covr test traces; [rd2markdown](https://github.com/Genentech/rd2markdown) for converting .Rd files into Markdown.

Also on reverse dependencies checks, Pawel Rucki and André Veríssimo presented
 [{verdepcheck} - A Tool for Dependencies Check](https://userconf2024.sched.com/event/1cC1c/verdepcheck-a-tool-for-dependencies-check-pawel-rucki-andre-verissimo-roche?iframe=no&w=100%&sidebar=yes&bg=no). [Slides](https://static.sched.com/hosted_files/userconf2024/2d/%7Bverdepcheck%7D%20-%20A%20Tool%20for%20Dependencies%20Check.pdf), [verdepcheck](https://insightsengineering.github.io/verdepcheck/main/).
 
Franciszek Walkowiak's topic was [Systems Integration Tests for R Package Cohorts](https://userconf2024.sched.com/event/1c8yp/systems-integration-tests-for-r-package-cohorts-franciszek-walkowiak-roche?iframe=no&w=100%&sidebar=yes&bg=no), including the introductions to two open-source utilities, [scribe](https://github.com/insightsengineering/scribe) that creates complete build, check and install reports for a collection of R projects and [locksmith](https://github.com/insightsengineering/locksmith) that helps with `renv.lock` creation. [Slides](https://static.sched.com/hosted_files/userconf2024/1c/systems-integration-tests-for-r-package-cohorts.pdf).

#### On good practice

Daniel Sabanés Bové introduced openstatsware's work on _minimal viable_ good practice standards for R packages, in an [online talk](https://virtualuser2024.sched.com/event/1eBda/openstatsguide-minimum-viable-good-practices-for-high-quality-statistical-software-packages-daniel-sabanes-bove-rconis?iframe=no&w=&sidebar=yes&bg=no).

Pedro Silva listed [Seven Deadly Sins Holding You
Back as a Software Developer](https://userconf2024.sched.com/event/1c8zm/seven-deadly-sins-holding-you-back-as-a-software-developer-pedro-silva?iframe=no&w=100%&sidebar=yes&bg=no). [Slides](https://static.sched.com/hosted_files/userconf2024/86/Seven%20Deadly%20Sins%20Holding%20You%20Back%20as%20a%20Software%20Developer.pdf).

Hugo Gruson had a poster on [A reproducible analysis of CRAN Task Views to understand the state of an R package ecosystem](https://userconf2024.sched.com/event/1cDry/poster-session-guided-tours-a-reproducible-analysis-of-cran-task-views-to-understand-the-state-of-an-r-package-ecosystem-hugo-gruson-dataorg?iframe=no&w=100%&sidebar=yes&bg=no). Find the [live analysis](https://epiverse-connect.github.io/ctv-analysis/).

#### On package design

Hugo Gruson highlighted the benefits of using S3 classes for [interoperability in Existing Software Ecosystems](https://userconf2024.sched.com/event/1c8z3/building-interoperability-in-existing-software-ecosystems-with-s3-classes-hugo-gruson-dataorg?iframe=no). [Slides](https://static.sched.com/hosted_files/userconf2024/ff/2024-07-09_Gruson%20-%20Building%20Interoperability%20with%20S3%20Classes.pdf).
See also his recent [blog posts on the topic](https://epiverse-trace.github.io/blog.html#category=S3).

Ligia Adamska used an onion analogy to explain [Layered Design for R Package
Development: Meeting the Needs of
Pharmaceutical R&D Stakeholders](https://userconf2024.sched.com/event/1c8z0/layered-design-for-r-package-development-meeting-the-needs-of-pharmaceutical-rd-stakeholders-ligia-adamska-msd-switzerland?iframe=no&w=100%&sidebar=yes&bg=no). [Slides](https://static.sched.com/hosted_files/userconf2024/7a/UseR_presentation_Final.pdf).

#### On tools

Daphne Grasselly, Franciszek Walkowiak and Pawel Rucki lead a tutorial on [Streamlining R package
development with GitHub Actions
Workflows](https://userconf2024.sched.com/event/1c8yI/tutorial-streamlining-r-package-development-with-github-actions-workflows-daphne-grasselly-pawel-rucki-roche-dinakar-kulkarni-genentech-pre-registration-required?iframe=no) whose [slides](https://static.sched.com/hosted_files/userconf2024/19/streamlining-r-package-development-with-github-actions-workflows.pdf) are available.

Emil Hvitfeldt explained how to [make better error messages with rlang and cli](https://virtualuser2024.sched.com/event/1eBdO/making-better-error-messages-with-rlang-and-cli-emil-hvitfeldt-posit-pbc?iframe=no&w=&sidebar=yes&bg=no).

Ella Kaye shared her insights on [C for R users](https://userconf2024.sched.com/event/1c8zS/c-for-r-users-ella-kaye-university-of-warwick?iframe=no&w=100%&sidebar=yes&bg=no). [Slides](https://static.sched.com/hosted_files/userconf2024/84/c-for-r-users.pdf).

Davis Vaughan introduced [tree-sitter, an efficient incremental parsing library] (https://userconf2024.sched.com/event/1c90k/improving-development-tooling-with-an-r-grammar-for-tree-sitter-davis-vaughan-posit?iframe=no&w=100%&sidebar=yes&bg=no). [R package with bindings to tree-sitter](https://davisvaughan.github.io/r-tree-sitter/) whose README states "tree-sitter is useful for a number of things, including syntax highlighting, go-to definition, code reshaping, and more."

#### On multilingualism

Elio Campitelli spoke about [Building Bilingual Bridges
with Multilingual Manuals](https://userconf2024.sched.com/event/1c8yj/building-bilingual-bridges-with-multilingual-manuals-elio-campitelli-universidad-de-buenos-aires?iframe=no&w=100%&sidebar=yes&bg=no). [Slides](https://static.sched.com/hosted_files/userconf2024/3d/UseR2024-1.pdf). See also their [post on our blog](https://ropensci.org/blog/2024/07/09/multilingual-documentation-wg/)!

#### On debugging

Shannon Pileggi delivered a tutorial on [debugging in R](https://events.linuxfoundation.org/user/program/schedule/). [Materials](https://github.com/rstats-wtf/wtf-debugging).

Antoine Fabri gave an overview of the motivations behind, and features of, his [constructive package](https://userconf2024.sched.com/event/1c8ym/constructive-a-nicer-dput-using-idiomatic-constructors-antoine-fabri-cynkra?iframe=no&w=100%&sidebar=yes&bg=no), that among other things can be useful for troubleshooting. [constructive](https://cynkra.github.io/constructive/).

#### On a last resort for archived CRAN packages

Henrik Bengtsson and Lluís Revilla had a poster about their [CRANhaven project](https://www.cranhaven.org/) that is a backup solution for end-users when a package falls of CRAN (and that is built using R-universe!).

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
