---
title: Introducing BaseSet
author:
  - Lluís Revilla Sancho
date: '2021-01-14'
slug: introducing-baseset
categories: []
tags:
  - BaseSet
  - Software Peer Review
  - data-munging
  - R
  - packages
  - community
package_version: 0.0.14
description: ""
twitterImg: blog/2021/01/14/introducing-baseset/name-of-image.png
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "", collapse = TRUE)

# knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)

# knitr hook to use Hugo highlighting options
knitr::knit_hooks$set(
  source = function(x, options) {
  hlopts <- options$hlopts
    paste0(
      "```r ",
      if (!is.null(hlopts)) {
      paste0("{",
        glue::glue_collapse(
          glue::glue('{names(hlopts)}={hlopts}'),
          sep = ","
        ), "}"
        )
      },
      "\n", glue::glue_collapse(x, sep = "\n"), "\n```\n"
    )
  }
)
```

The figure paths in the resulting Markdown will already be valid so you won't need to edit the Markdown file.

## Introducing sets

[Brief history](https://b101nfo.blogspot.com/2020/07/my-journey-on-bioconductor.html)

When one `join_*` or `merge` two tables or you do an `intersection` and `setdiff` between two vectors you are using set operations coming from [set theory](https://en.wikipedia.org/wiki/Set_theory).
Sets are closed groups of elements that have been studied for many years and have many utilities.

Citation of the primary literature[^1].

[^1]: Sciaini, M., Fritsch, M., Scherer, C., & Simpkins, C. E.
    (2018).
    NLMR and landscapetools: An integrated environment for simulating and modifying neutral landscape models in R.
    Methods in Ecology and Evolution, 9(11), 2240-2248.
    <https://doi.org/10.1111/2041-210X.13076>

Citation of a website[^2].

[^2]: Elin Waring, Michael Quinn, Amelia McNamara, Eduardo Arino de la Rubia, Hao Zhu and Shannon Ellis (2019).
    skimr: Compact and Flexible Summaries of Data.
    R package version 1.0.7.
    <https://CRAN.R-project.org/package=skimr>

Citation of an R package[^3].

[^3]: Hugo static site generator.
    <https://gohugo.io/>

### Subsection heading

We recommend the use of [Hugo shortcodes](https://gohugo.io/content-management/shortcodes/) to include images, tweets, videos, gists, etc.
In the Rmd file they need to be between `<!--html_preserve-->` and `<!--/html_preserve-->`.

**Add an image** by using a Hugo shortcode.
The image is saved under `/content/blog/YYYY-MM-DD-slug/name-of-image.png`.

<!--html_preserve-->
{{< figure src = "name-of-image.png" width = "400" alt = "this is the alternative text" >}}
<!--/html_preserve-->

Consult the Technical Guidelines for tips on changing image size, alignment, and for advice on alternative text.

Below is another image, a plot generated by a code chunk, so you might see how to add alternative text and other options in this case.
You can also use [all options documented for the figure shortcode](https://gohugo.io/content-management/shortcodes/#figure) except for `src` that will be created automatically when knitting.

```{r chunkname, hugoopts=list(alt="alternative text please make it informative", caption="this is what this image shows, write it here or in the paragraph after the image as you prefer", width=300)}
plot(1:10)
```

Once this file is knitted the plot above will be inserted with the correct syntax.

**Embed a tweet** by using a Hugo shortcode.

<!--html_preserve-->
{{< tweet 1138216112808529920 >}}
<!--/html_preserve-->

**Tweak code highlighting** Below you can see how to use Hugo [Chroma code highlighting options](https://gohugo.io/content-management/syntax-highlighting/#highlight-shortcode): highlight the 1st and 4th to 6th code lines, number lines as a table starting from 199.

```{r hl, hlopts = list(linenos='table',hl_lines='[1,"4-6"]',linenostart=199)}
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
a <- 1
```

**Add citation or footnote** text by using the format below
