---
title: How to Update a Translation with Babeldown
author: 
- Maëlle Salmon
date: '2024-02-09'
slug: deepl-update-babeldown
categories: []
output: hugodown::md_document
tags:
  - tech notes
  - multilingual
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  comment = ''
)
```
As part of our [multilingual publishing project](/multilingual-publishing/), and with [funding from the R Consortium](https://www.r-consortium.org/all-projects/awarded-projects/2022-group-2), we've worked on the R package [babeldown](https://docs.ropensci.org/babeldown/) for translating Markdown-based content using the DeepL API.
In this tech note, we'll show how you can use babeldown to update a translation after you've slightly edited a document in its main language version!

## Initial situation: an English document and its French translation

Let's assume we have an English document called `bla.md`.

```{r}
dir <- withr::local_tempdir()
file <- file.path(dir, "bla.md")
fs::file_create(file)
english_text <- c("# header", "", "this is some text", "", "## subtitle", "", "nice!")
brio::write_lines(english_text, file)
```

```{r, echo=FALSE}
cat(paste(english_text, collapse = "\n"))
```

We'd have translated it with babeldown, then edited the translation manually.

```{r, eval=FALSE}
Sys.setenv("DEEPL_API_URL" = "https://api.deepl.com")
Sys.setenv(DEEPL_API_KEY = keyring::key_get("deepl"))

out_file <- file.path(dir, "bla.fr.md")
deepl_translate(
  path = file,
  out_path = out_file,
  source_lang = "EN",
  target_lang = "FR",
  formality = "less",
  yaml_fields = NULL
)
```

```{r, echo=FALSE}
out_file <- file.path(dir, "bla.fr.md")
french_text <- c("# titre", "", "ceci est du texte", "", "## sous-titre", "", "chouette !")
brio::write_lines(french_text, out_file)
```

Here's the French text:


```{r, echo=FALSE}
cat(paste(french_text, collapse = "\n"))
```

At this stage let's set up the Git infrastructure for the folder containing the two documents.
In real life, we might already have it in place.
The important thing is to start tracking changes before we edit the English document again.

```{r}
gert::git_init(dir)
gert::git_config_set("user.name", "Jane Doe", repo = dir)
gert::git_config_set("user.email", "jane@example.com", repo = dir)
gert::git_add(c(fs::path_file(file), fs::path_file(out_file)), repo = dir)
gert::git_commit_all("First commit", repo = dir)
gert::git_log(repo = dir)
```

## Changing the English document

Now imagine we change the English document.

```{r}
brio::write_lines(
  c("# a title", "", "this is some text", "", "awesome", "", "## subtitle", ""),
  file
)
gert::git_add(fs::path_file(file), repo = dir)
gert::git_commit("Second commit", repo = dir)
```