---
title: How to Update a Translation with Babeldown
author: 
- Maëlle Salmon
date: '2024-02-09'
slug: deepl-update-babeldown
categories: []
output: hugodown::md_document
tags:
  - tech notes
  - multilingual
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  comment = ''
)
```
As part of our [multilingual publishing project](/multilingual-publishing/), and with [funding from the R Consortium](https://www.r-consortium.org/all-projects/awarded-projects/2022-group-2), we've worked on the R package [babeldown](https://docs.ropensci.org/babeldown/) for translating Markdown-based content using the DeepL API.

In a previous tech note we exemplified the [use of babeldown for translating a blog post](/blog/2023/09/26/how-to-translate-a-hugo-blog-post-with-babeldown/), in a workflow supported by Git.
Now, we use babeldown for translating **living** documents, such as our regularly updated dev guide, where translations do not only need to be created but also kept in sync with the main language version.

In this tech note, we'll show how you can use babeldown to update a translation after you've slightly edited a document in its main language version!

## Initial situation: an English document and its French translation

Let's assume we have an English document called `bla.md`.

```{r}
dir <- withr::local_tempdir()
file <- file.path(dir, "bla.md")
fs::file_create(file)
english_text <- c("# header", "", "this is some text", "", "## subtitle", "", "nice!")
brio::write_lines(english_text, file)
```

```{r, echo=FALSE}
cat(paste(english_text, collapse = "\n"))
```

We'd have translated it with babeldown, then edited the translation manually.

```{r, eval=FALSE}
Sys.setenv("DEEPL_API_URL" = "https://api.deepl.com")
Sys.setenv(DEEPL_API_KEY = keyring::key_get("deepl"))

out_file <- file.path(dir, "bla.fr.md")
deepl_translate(
  path = file,
  out_path = out_file,
  source_lang = "EN",
  target_lang = "FR",
  formality = "less",
  yaml_fields = NULL
)
```

```{r, echo=FALSE}
out_file <- file.path(dir, "bla.fr.md")
french_text <- c("# titre", "", "ceci est du texte", "", "## sous-titre", "", "chouette !")
brio::write_lines(french_text, out_file)
```

Here's the French text:


```{r, echo=FALSE}
cat(paste(french_text, collapse = "\n"))
```

At this stage let's set up the Git infrastructure for the folder containing the two documents.
In real life, we might already have it in place.
The important thing is to start tracking changes before we edit the English document again.

```{r}
gert::git_init(dir)
gert::git_config_set("user.name", "Jane Doe", repo = dir)
gert::git_config_set("user.email", "jane@example.com", repo = dir)
gert::git_add(c(fs::path_file(file), fs::path_file(out_file)), repo = dir)
gert::git_commit_all("First commit", repo = dir)
gert::git_log(repo = dir)
```

## Changing the English document

Now imagine we change the English document.

```{r}
brio::write_lines(
  c("# a title", "", "this is some text", "", "awesome", "", "## subtitle", ""),
  file
)
gert::git_add(fs::path_file(file), repo = dir)
gert::git_commit("Second commit", repo = dir)
```

## Updating the translation

We don't want to send the whole document to DeepL API again!
Indeed, the translation of text that hasn't changed shouldn't change: that translation results from careful work by human translators. 
Furthermore, if we were to send it to the API again, we'd use credits again.

Fortunately we have two babeldown functions at our disposal:

- `babeldown::deepl_translate_markdown_string()` in case we can easily see what fragment of text has bee changed. 
- `babeldown::deepl_update()` that operates more automatically, but might change more text than needed as it will send a whole paragraph to DeepL API if it changed, as opposed to a single sentence.

## How `babeldown::deepl_update()` works under the hood

In case you're curious, `babeldown::deepl_update()` doesn't use the Git diff at all!
Although that definitely was the first idea we explored.

`babeldown::deepl_update()` does scour the Git log to use the snapshot of the main language version that was in sync with the translation.
It's the "old English document", that goes with the "old French document".
We have the "new English document" and what's missing is the "new French document".
We want that new French document to use as much as possible of the old French document, only using an automatic translation for the parts that are new.

The function uses an XML representation of the documents, as created by [tinkr](https://docs.ropensci.org/tinkr).
A necessary condition for using `babeldown::deepl_update()` is that the old English document and the old French document need to have the same XML structure: say, one heading followed by two paragraphs then a list.

For each child of the body of the new English document (a paragraph, a list, a heading...), `babeldown::deepl_update()` tries finding the same tag in the old English document ([identified by having the same `xml2::xml_text()` and children of the same type](https://github.com/ropensci-review-tools/babeldown/blob/e8ef563eca232f1ba08b84c7bb0f999d4f452938/R/update.R#L136)).
If it finds the same tag, it uses the tag located at the same position in the old French document.
If it does not find it, it sends it to DeepL API.

## Conclusion

In this post we explained how to use babeldown to update translations of living document.
We at rOpenSci are ourselves users of babeldown for this scenario!
Maintaining translations is time consuming but important work.