---
title: rOpenSci News Digest, May 2021
author:
  - The rOpenSci Team
date: '2021-05-21'
slug: ropensci-news-digest-may-2021
categories: []
tags: 
  - newsletter
description: social co-working, R-universe, rOpenSci at useR! 2021, new packages and package news
output:
  html_document:
    keep_md: yes
params:
  last_newsletter: '2021-04-16'
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

We have no less than three categories of (online :wink:) events coming up: social co-working events, a community call, a conference with contributions from four rOpenSci staff members. :sparkles:

### Community Calls

* Our next community call is about [**rOpenSci's R-universe Project**](/commcalls/may2021-r-universe/) and will take place on Tuesday, 25 May 2021 17:00 UTC. The R-universe platform is a new umbrella project under which rOpenSci experiments with new ideas for improving publication and discovery of research software packages in R. In this 1-hour community call, **Jeroen Ooms** will explain the basic steps of setting up your own universe, and getting started with publishing packages (including experimental software, development versions, [research compendia](https://ropensci.org/blog/2021/04/09/r-universe-articles/)) and articles on your [personal subdomain](https://jeroen.r-universe.dev). We are interested in hearing your thoughts and ideas on how this infrastructure can improve how researchers publish and discover R packages and related content.

* Video and resources from our past community call [**Set Up Your Package to Foster a Community**](/commcalls/apr2021-pkg-community/) were posted. rOpenSci puts ongoing effort into exploring and communicating how developers can best attract attention to their package (e.g. usage, citations, or feedback), or how to set up their repository to encourage the types of contributions they want. In this 1-hour community call, **Maëlle Salmon**, **Hugo Gruson**, and **Steffi LaZerte** shared tips and examples on how to do this! **Sina Rüeger** [summarized the call](/blog/2021/04/28/commcall-pkg-community/) with help from Mark Padgham.

### Social Co-Working

We held our first two **Social Co-Working and Label-athon events** on April 29th and May 13th. 
These were truly **international events** with 14 participants from Canada, Germany, Switzerland, USA, the UK, France, Singapore, and Mexico. 
We had a blast **socializing (scavenger hunts!) and co-working on our personal R projects** including repository organization, creating/labelling issues, preliminary research, personal GitHub README files, and those packages you can never find time for. 
We had **quiet and noisy rooms** to work in, so everyone was able to find an environment that worked best for them.
Sound like fun? We have **two more coming up**! 

- [Thursday May 27th 9am Pacific](/events/labelathon03/) (16:00 UTC) - [Registration encouraged!](https://form.jotform.com/211236537172249)

- [Thursday June 10th 5pm Pacific](/events/labelathon04/) (Fri June 11th 00:00 UTC) - [Registration encouraged!](https://form.jotform.com/211236201763243)

### rOpenSci at useR! 2021

[Registration](https://user2021.r-project.org/participation/registration/) for the **useR! 2021 conference** has begun, with  Early Bird Deadline until May 28th. 
useR! conferences are non-profit conferences organized by community volunteers for the community, supported by the R Foundation.
useR! 2021 will take place online.

Get excited for four contributions by rOpenSci staff members:

* **Jeroen Ooms**, rOpenSci Lead Infrastructure Engineer, will give a [**keynote talk**](https://user2021.r-project.org/program/keynotes/) about **building a modern and scalable package build infrastructure**, such as R-universe.

* **Maëlle Salmon**, Research Software Engineer with rOpenSci, will give a [tutorial](https://user2021.r-project.org/program/tutorials/), _**GET better at testing your R package!**_ on Wednesday, July the 7th at 9-12AM GMT+2. This tutorial is about advanced testing of R packages, with HTTP testing as a case study.

* **Stefanie Butland**, rOpenSci community manager, will present on **rOpenSci's Model for Managing a Federated Open Source Software Community**.

* **Mark Padgham**, rOpenSci Software Research Scientist, will also give a regular talk, about his original **autotest** package for automatic testing of packages.

Find out about more [events](/events).

### Software review in Spanish

Beside events, another great piece of news: censo2017 was the first package to pass **rOpenSci software peer-review in Spanish**! :tada:
The [censo2017](https://github.com/ropensci/software-review/issues/414) package by [Pachá (aka Mauricio Vargas Sepúlveda)](/author/pachá-aka-mauricio-vargas-sepúlveda/) provides access to the 2017 Chilean Census.
Editor [Melina Vidoni](/blog/2019/01/31/more_editors/) was in charge of this review, the reviewers were [Frans van Dunné](https://github.com/FvD) and [María Paula Caldas](https://www.mpaulacaldas.com/).
Frans van Dunné most kindly translated our [review template to Spanish](https://devguide.ropensci.org/reviewtemplatees.html)!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", toString(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }

  if (nzchar(package$onboarding) || package$package == "censo2017") {
    if (package$package == "censo2017") {
      rev <- c("Frans van Dunné", "Maria Paula Caldas")
      package$onboarding <- "https://github.com/ropensci/software-review/issues/414"
    }
    if (package$package == "exoplanets") {
      rev <- c(rev, "Nichole Monhait")
    }
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", toString(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- toString(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the latest newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue)
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s.", name, url, description, user))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  yaml <- rmarkdown::yaml_front_matter(path)
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = toString(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    twitterImg = yaml$twitterImg %||% "",
    twitterAlt = yaml$twitterAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug
    )

  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, seq(nrow(posts)))
format_post <- function(post) {
  url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(post$date),
    stringr::str_pad(lubridate::month(post$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(post$date), 2, "0", side = "left"),
    post$slug
    )
  string <- sprintf("* [%s](%s) by %s", post$title, url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$twitterImg != "") {
    img_file <- fs::path_file(post$twitterImg)
    download.file(sprintf("https://ropensci.org/%s", post$twitterImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$twitterAlt)
    )
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, "software_peer_review")]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, "other")]
if (length(others) > 0) {
  cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, "tech_note")]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Citations

```{r cit}
citations <- jsonlite::read_json("https://ropensci-org.github.io/ropensci_citations/citations_all_parts_clean.json")
```

No new citations added to our database this month ([browse all citations](/citations)).
Thanks for citing our tools!

```{r citations, results = "asis", eval=FALSE}
since <- lubridate::as_date(last_newsletter) - 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci-org",
  repo = "ropensci_citations",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  )
)
old_commit <- gh::gh("/repos/{owner}/{repo}/commits/{ref}",
  owner = "ropensci-org",
  repo = "ropensci_citations",
  ref = commits[[length(commits)]]$parents[[1]]$sha
  )
old <- "https://raw.githubusercontent.com/ropensci-org/ropensci_citations/%s/citations_all_parts_clean.json" %>%
    sprintf(old_commit$sha) %>%
    jsonlite::read_json() 

new <- citations[length(old):length(citations)]

format_package <- function(package, packages = packages) {
  if (package %in% packages) {
    package <- sprintf("[**%s**](https://docs.ropensci.org/%s)", package, package)
  } else {
    package <- sprintf("**%s**", package)
  }
  package
}

format_one <- function(citation, packages) {
  packages <- toString(purrr::map_chr(citation$name, format_package, packages = packages))
  sprintf("* %s in %s\n\n", packages, citation$citation)
}

cat(
  paste0(
    sort(purrr::map_chr(new, format_one, packages = registry$package)),
    collapse = ""
  )
)

```

## Use cases

```{r usecases}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
usecases <- dplyr::filter(usecases, date >= as.Date(last_newsletter))
usecases <- split(usecases, seq(nrow(usecases)))
```

`r snakecase::to_sentence_case(english(length(usecases)))` use cases of our packages and resources have been reported since we sent the last newsletter.

```{r usecases2, results='asis'}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", usecase$title, usecase$url, usecase$reporter)
  string
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)!

## Call for maintainers

There's no open call for new maintainers at this point but you can refer to our [contributing guide](https://contributing.ropensci.org/) for finding ways to get involved!
As the maintainer of an rOpenSci package, feel free to contact us on Slack or email `info@ropensci.org` to get your call for maintainer featured in the next newsletter. 

## Package development corner

Some useful tips for R package developers. :eyes:

Sometimes you find yourself documenting internal functions with roxygen2 without really wanting everyone to get to see the docs.
Do you know the [**difference between the `@internal` and `@NoRd` tags**](https://community.rstudio.com/t/keywords-internal-vs-nord/35119/3)?
* Keyword [`@internal`](https://roxygen2.r-lib.org/articles/rd.html#indexing) means there will be a manual page for the function but it won't be in the documentation index. " `@keywords internal` is useful because it removes the function from the documentation index; it’s useful for functions aimed primarily at other developers, not typical users of the package."
* `@NoRd` means there won't even be a manual page so the roxygen2 documentation of that function only lives in the R script. 
We at rOpenSci [recommend not having manual pages for internal functions](https://devguide.ropensci.org/building.html): " Add `#' @noRd` to internal functions". It is also recommended in the [tidyverse style guide](https://style.tidyverse.org/documentation.html#internal-functions).

Speaking of documentation, when writing **examples**, do you know the difference between **`\dontrun{}`**, **`\donttest{}`** and **`\dontshow{}`**?
(Is this newsletter section turning into a quizz? :sweat_smile:)
The roxygen2 docs feature [a  table](https://roxygen2.r-lib.org/articles/rd.html#functions) explaining when different types of examples are run depending on the context (in `example()`,	`help()`, `R CMD check`, `R CMD check –as-cran`) which is very useful when you wonder how to safeguard your package examples on e.g. CRAN.
The same documentation section introduces the very handy **`@examplesIf`** roxygen2 tag that you can use for finer control.

Now on to **code archaeology**!
Do you want to ["pinpoint the version at which a certain argument appeared in a specific function of a package (or that its default value changed)"](https://twitter.com/JennyBryan/status/1343636610982662144), e.g. to figure out the minimal version required?
You might like the [**apicheck package** by David Hugh-Jones](https://github.com/hughjonesd/apicheck), that helps you "explore the historical API of functions in CRAN packages"; and [**rcheology** by the same author](https://github.com/hughjonesd/rcheology) for data on base packages.
Related to this are:

* [Using **git blame on the package NAMESPACE** to see when a function appeared](https://twitter.com/JennyBryan/status/1387439721157038080),
* The R-hub's [**CRAN source code mirror**](https://docs.r-hub.io/#cran-source-code-mirror): each repo is a CRAN package, each commit is the change between two versions of that package.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Twitter account](https://twitter.com/ropensci).
