---
title: 'rOpenSci News Digest, May 2024'
author: 'The rOpenSci Team'
date: '2024-05-24'
tags:
  - newsletter
slug: ropensci-news-may-2024
description: Champions program update, metadata, new packages and package news
params:
  last_newsletter: '2024-04-22'
---


```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ
### rOpenSci in the Research Organization Registry

rOpenSci was added to the [Research Organization Registry](https://ror.org/) (ROR) in its latest [release](https://github.com/ror-community/ror-updates/releases/tag/v1.46). The ROR is  a global, community-led registry of open persistent identifiers for research organizations. ROR IDs help link and disambiguate metadata about organizations in the scholarly record, much like DOIs and ORCiDs do for manuscripts and researchers. [Linked metadata is rOpenSci's love language ❤️ !](/blog/2018/03/16/thanking-reviewers-in-metadata/) Find us at <https://ror.org/019jywm96>.

### What are the rOpenSci Champions up to now?

Our Champions and mentors have been carrying out various programmed activities.
The first stage of the program has a very important training component. This year, we divided the training into technical and community tracks, with several one to two-hour sessions each. Here, you can find the openly available material for each session.

The Technical Track is focused on good software and package development practices: 

* [Beautiful Code, Because We’re Worth It!](https://stylish-code.netlify.app/#/): a 1-hour session on [good practices for  writing code](/blog/2024/02/22/beautiful-code/). 

* [Package Development: The Mechanics](https://rpkgdev-mechanics-2024.netlify.app/): three 2-hour sessions for beginners to demystify the creation of an R package.

* [Package Development: Not Rocket Science](https://rpkgdev-rocket-2024.netlify.app/): a 2-hour hands-on workshop for advanced R package development with tips and tricks. 

* [How rOpenSci Performs Peer Review](https://github.com/maurolepore/ropensci-review): a 2-hour workshop to showcase and discuss how rOpenSci communicates, builds, and reviews software. 

The Community Track is focused on community-building skills:

* [How To Contribute to Open Projects and Communities](https://osschampionsprogram.netlify.app/): a 2-hour session presenting community participation frameworks, examples, and recommendations for your package. 

* [Marketing Ideas For Your Package](/blog/2024/03/07/package-marketing/): A blog post presenting a series of activities and tools for advertising your package.

Next month, we will complete the training with a workshop on Git and GitHub and another on Event Organization. 

### Coworking

Read [all about coworking](/blog/2023/06/21/coworking/)!

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.

- [Tuesday, June 4th, 09:00 Australia Western (01:00 UTC)](/events/coworking-2024-06), R in the Wild with cohosts [Ernest Guevarra](/author/ernest-guevarra/), [Tomás Zaba](/author/tom%C3%A1s-zaba/), [Nicholus Tint Zaw](/author/nicholus-tint-zaw/), [Zython Paul Lachica](/author/zython-paul-lachica/) and [Steffi LaZerte](/author/steffi-lazerte).
    - Work on some of your own wild workflows
    - Learn how others use R in the Wild (at work, with or without teams)
    - Chat with our cohosts, discuss the challenges they face, and wins they have through their work

- [Tuesday, July 2nd, 14:00 Europe Central (12:00 UTC)](/events/coworking-2024-07), Git and GitHub with cohost [Zhian Kamvar](/author/zhian-n.-kamvar/) and [Steffi LaZerte](/author/steffi-lazerte).
    - Read up on Git and GitHub and how they might serve you
    - Start your first Git/GitHub project
    - Chat with our cohost about the pros and cons of Git & GitHub, resources for getting started and tips and tricks.

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

## Software :package:


### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% c("tabulapdf", old$package),
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    sub(
      "<doi:",
      "<https://doi.org/",
      sub(
      "standarizes",
      "standardizes",
      sub("\\.$", "", stringr::str_squish(package$details))
    )
    )
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- 
    sub(
      "\\. \\. ",
      ". ",
      paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    )
    if (length(rev) > 0) {
      
   
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
      
    } else {
       if (package$package == "wmm") {
      string <- paste0(string, " by Athanasia Monika Mowinckel and Rohit Goswami.")
       } else if (package$package == "birdsize") {
      string <- paste0(string, " by Quentin Read and Matt Strimas-Mackey.")
    } else if (package$package == "naijR") {
      string <- paste0(string, " by Margaret Siple and Alican Cagri Gokcek.")
    } else {
      string <- paste0(string, ".")
    }
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
packages[purrr::map_chr(packages, "package") == "gert"][[1]]$owner <- "r-lib"
packages[purrr::map_chr(packages, "package") == "gert"][[1]]$repo <- "gert"
packages[purrr::map_chr(packages, "package") == "credentials"][[1]]$owner <- "r-lib"
packages[purrr::map_chr(packages, "package") == "credentials"][[1]]$repo <- "credentials"
packages[purrr::map_chr(packages, "package") == "commonmark"][[1]]$owner <- "r-lib"
packages[purrr::map_chr(packages, "package") == "commonmark"][[1]]$repo <- "commonmark"
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  if (url == "https://wwww.researchgate.net/profile/Kristof_Haneca") url <- "https://www.researchgate.net/profile/Kristof_Haneca"
  if (url == "https://www.surfaces.co.il") url <- "https://github.com/micha-silver"
  if (url == "http://csaybar.github.io") url <- "https://csaybar.github.io"
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }
  software_peer_review <- "Software Peer Review" %in% yaml$tags
  tech_note <- "tech notes" %in% yaml$tags && !software_peer_review
  other <- !software_peer_review && !tech_note
  
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = software_peer_review,
    tech_note = tech_note,
    other = other,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = stringr::str_squish(yaml$description %||% ""),
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    if (post$title != "R-universe now builds MacOS ARM64 binaries for use on Apple Silicon (aka M1/M2/M3) systems") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
    }
  } else {
    string <- paste0(string, ".")  
  }
    if (post$title == "Marketing Ideas For Your Package") {
      string <- paste0(
      string,
      " This post was discussed on the [R Weekly highlights podcast](https://rweekly.fireside.fm/156) hosted by Eric Nantz and Mike Thomas.")
  }
  if (post$title == "Beautiful Code, Because We’re Worth It!") {
    string <- paste0(string, " This post was featured in the [R Weekly Highlights podcast](https://rweekly.fireside.fm/154?t=1338) hosted by Eric Nantz and Mike Thomas.")
  }
  
  if (post$socialImg != "") {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt)
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

* [historydata](https://docs.ropensci.org/historydata/), datasets for historians. [Issue for volunteering](https://github.com/ropensci/historydata/issues/23).

* [USAboundaries](https://docs.ropensci.org/USAboundaries/) (and USAboundariesdata), historical and contemporary boundaries of the United States of America . [Issue for volunteering](https://github.com/ropensci/USAboundaries/issues/50).

### Calls for contributions

* [Help make qualtRics better! Code contributions wanted](/blog/2024/03/01/qualtrics-call-help/)

* [Help make assertr better! Come close issues](/blog/2024/02/27/assertr-call-help/)

* [Help users and developers of targets by answering questions!](/blog/2024/02/29/targets-call-help/)

* [Help make waywiser better! User requests wanted](/blog/2024/02/19/waywiser-call-help/)

Also refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.

## Package development corner

Some useful tips for R package developers. :eyes:

### Token now needed for code coverage via codecov and {covr} on GitHub Actions

[Test coverage](https://r-pkgs.org/testing-design.html#sec-testing-design-coverage) reports are useful when assessing and improving tests of an R package.
One can run the covr package locally, or send results to an interface like codecov.io that provides interactive exploration of the output.
It is possible to compute test coverage and send the results to codecov.io on GitHub Actions, using the r-lib/actions actions.

Now, recently, workflows have started to fail if one did not set a codecov.io token as an environment variable. 
How to solve this

- For an individual repo, add [your codecov.io token as `CODECOV_TOKEN` to the repository secrets](https://docs.codecov.com/docs/adding-the-codecov-token).
- For an organization, ask the organization owner to create an organization wide codecov.io token and to store it as an organization secret. Make sure to update your [GitHub Actions workflow file](https://github.com/r-lib/actions/issues/834#issuecomment-2102337428), for instance by running [`usethis::use_github_action("test-coverage")`](https://usethis.r-lib.org/reference/use_github_action.html).

### Notes on package repositories

Lluís Revilla Sancho wrote about [Packaging R: getting in repositories](https://llrs.dev/post/2024/05/05/packaging-r-getting-in/).
We particularly note his two definitions of package repositories: the first one consists in making `install.packages()` work, the second in adding a layer of checks to packages in the repository. 
He then added "[R-universe](/r-universe/) is using the first definition but could be used to generate repositories with checks that comply with the second definition."

### Standardize a DESCRIPTION file in one function call

Do you know about the [`desc::desc_normalize()`](https://desc.r-lib.org/reference/desc_normalize.html) function that orders and formats DESCRIPTION fields in a standard way?
Once you start using it, there's no way back.

You can also call it indirectly via [`usethis::use_tidy_description()`](https://usethis.r-lib.org/reference/tidyverse.html) that also sets the Encoding field to UTF-8.

### Automated refactoring with xmlparsedata

Refactoring code can be tedious manual work, but it can also be tedious automated work. :wink:
Read a post about how to [replace all occurrences of a given function call with another one](https://masalmon.eu/2024/05/15/refactoring-xml/) using an XML representation of the code.

### What's your favorite IDE?

Athanasia Monica Mowinckel wrote an informative post about the [IDEs she uses](https://drmowinckels.io/blog/2024/ide/).

### Code review resources

Beside the [tidyverse code review guidance](https://code-review.tidyverse.org/), we can now recommend you check out the [code review anxiety workbook](https://developer-success-lab.gitbook.io/code-review-anxiety-workbook-1) by Carol Lee and Kristen Foster-Marks, that explains what code review anxiety is, and describes efficient methods to deal with it.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
