---
slug: "news-april-2025"
title: rOpenSci News Digest, April 2025
author:
  - The rOpenSci Team
date: '2025-04-28'
tags:
  - newsletter
description:
params:
  last_newsletter: '2025-03-27'
  doi: "10.59350/a6syn-6hc72"
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(trimws(...), sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### Event: Career Paths for U.S. Federal Data Scientists

rOpenSci and Openscapes are lucky to have as members many U.S. federal employees who enrich our work and community. In a time of great uncertainty for so many government workers, we want to support our members who are considering their next career steps.

Join us on Tuesday, May 6 at 8:30PM EDT for a discussion and networking event for and with current and former U.S. federal government data scientists.

[More information](/events/datasci-careers-feds/).

### rOpenSci Champions Program 2025 In Spanish: Apply before April 30th!

Reminder: The call for applications to be part of the new cohort of our 2025 Program is open until April 30th! 
For the first time it will be in Spanish.

Find out more in our [call for applications](/blog/2025/03/10/champeons-latin-america/) open until **Wednesday, April 30, 2025**.

### Stay safe online

Our executive director Noam Ross was interviewed in this Nature article ["Stay safe from online hate with these five tips"](https://archive.ph/pIZpx).

Another relevant article on the topic is this 2021 paper by Danielle Smalls and Greg Wilson, ["Ten quick tips for staying safe online"](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1008563).  

### Which programming language should I use?

Our community manager Yanina Bellini Saibene was also interviewed in Nature, for the article ["Which programming language should I use? A guide for early-career researchers"](https://archive.ph/cvGwW)

> “The user community is also a key consideration — Saibene says that’s one of the things she loves most about R. The language’s large, welcoming and engaged user community means that tools are developed and updated frequently; that local user groups exist around the world; and that tutorials and other resources are available in many spoken languages, including her native Spanish.”

    
### Coworking 

Read [all about coworking](/blog/2023/06/21/coworking/)!

- Tuesday May 6th 9:00 Australia Western (01:00 UTC), "Writing functions in R" with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Nicholas Tierney](/author/nicholas-tierney/).
    - Write some functions;
    - Explore new or interesting ways of writing functions in R.

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE,eval=TRUE}
packages <- split(new, seq(nrow(new)))
packages <- rev(packages)
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  
  if (package$package == "eDNAjoint") {
    desc_link <- "https://raw.githubusercontent.com/ropensci/eDNAjoint/26fb2d22dcd4314cdf7d170812c6247ae51f25a0/DESCRIPTION"
  }
  
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  if (package$package == "mbquartR") {
    rev <- c(rev, "Sheila Saia")
  }
  if (package$package == "geotargets") {
    rev <- c("Anthony Martinez", "Denisse Fierro Arcos")
  }
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (package$package == "fireexposuR") {
      rev <- c("Sherry Zhang ", "Ronny A. Hernández Mora")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state,
    .limit = Inf,
    .progress = FALSE
  )

  issues <- purrr::keep(issues, \(x) label %in% purrr::map_chr(x[["labels"]], "name"))
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- glue::glue('https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A%22{label}%22')
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))

  if (lines[2] == "headless: true") {
    return(NULL)
  }

  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md",  "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$socialImg != "") {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string, "\n```{=html}\n",
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt),
      "\n```\n"
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

* [photosearcher](https://docs.ropensci.org/photosearcher/), Searches Flickr for photographs and metadata. [Issue for volunteering](https://github.com/ropensci/photosearcher/issues/21).

* [MODIStsp](https://docs.ropensci.org/MODIStsp/), automatic download and preprocessing of MODIS Land Products Time Series. [Issue for volunteering](https://github.com/ropensci/MODIStsp/issues/262).

* [hddtools](https://docs.ropensci.org/hddtools/), Hydrological Data Discovery Tools. [Issue for volunteering](https://github.com/ropensci/hddtools/issues/36).

* [USAboundaries](https://docs.ropensci.org/USAboundaries/) (and USAboundariesdata), historical and contemporary boundaries of the United States of America . [Issue for volunteering](https://github.com/ropensci/USAboundaries/issues/50).

* [historydata](https://docs.ropensci.org/historydata/), datasets for historians. [Issue for volunteering](https://github.com/ropensci/historydata/issues/23).

### Calls for contributions

Refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.

## Package development corner

Some useful tips for R package developers. :eyes:

### Eleven quick tips for writing a Bioconductor package

If you are planning to publish your package with Bioconductor, do not miss this [paper](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1012856) by Charlotte Soneson and collaborators.


### Dump all package information into one file for LLMs

With the [rdocdump](https://www.ekotov.pro/rdocdump/) package by Egor Kotov, you can obtain a single file containing the source code, documentation and vignettes of an R package.
You can then use the file for ingestion into a Large Language Model (LLM).

### plumber2

If you use the plumber package to publish your code as an API, you'll be interested in its successor, the [plumber2](https://posit-dev.github.io/plumber2/) package by Thomas Lin Pedersen.

### Use a nested folder structure for your package

If you dream of using nested folders for storing the R code of an package, you can try out the experimental [dir](https://github.com/cynkra/dir) package by Antoine Fabri.

### Minimal type guessing

Did you know about [minty](https://gesistsa.github.io/minty/), an R package by Chung-hong Chan that shares the the type inferencing and parsing tools used by the readr package prior to 2021?
It can be handy for your parsing needs.

### Visualising R Package Risk Assessments using Litmus

An interesting [post and dashboard](https://www.jumpingrivers.com/blog/litmus-dashboard/) by Pedro Silva, Astrid Radermacher & Colin Gillespie to assess the risk of packages.

### CI: Pin the R version if using renv

A good tip by Christophe Dervieux in our [Slack workspace](https://contributing.ropensci.org/resources.html#channels): 
If you use renv to freeze dependencies in continuous integration, you should pin the R version to what's been used to build the renv's lockfile.
You should update (and check) everything at once, when at all: the R versions and package versions.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
