---
slug: "news-june-2026"
title: rOpenSci News Digest, June 2026
author:
  - The rOpenSci Team
date: '2026-06-29'
tags:
  - newsletter
description: 15 years of rOpenSci; champions program; maintainer month; new packages and package news
params:
  last_newsletter: '2026-05-28'
  doi: "10.59350/312sy-1dt83"
---

```{r setup, include=FALSE}
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
  "/blog/%s/%s/%s/%s",
  lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
  stringr::str_pad(
    lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    2,
    "0",
    side = "left"
  ),
  stringr::str_pad(
    lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    2,
    "0",
    side = "left"
  ),
  rmarkdown::yaml_front_matter(knitr::current_input())$slug
)
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(trimws(...), sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ


### Champions Program update

We have two concurrent cohorts, both in Spanish. 

The 2025–2026 cohort is nearing the end of its participation in the program, so we are organizing the closing meeting and the overall evaluation.

The 2026--2027 cohort is continuing their training activities, meeting with their mentors, and starting to work on their packages, and they have been formally introduced on our blog! 
Read all about the [11 new champions](/blog/2026/06/09/champions-2026/).


### New editors Ronny Hernandez Mora, Joel Nitta, and Nick Tierney

We're thrilled to [introduce](/blog/2026/06/11/neweditorsq22026/) new editors Ronny Hernandez Mora, Joel Nitta, and Nick Tierney.
An official welcome and thank you to the three of them!

### A New Guide: Organizing Events for First-time Contributors

Steffi LaZerte and Yanina Bellini Saibene released a fantastic new rOpenSci guide!
Learn how to organize events for first-time contributors such as mini-hackathons and mini-translathons.
Read more in the [release announcement](/blog/2026/06/02/ftc-guide/).

### R-universe updates

"Five recent R-universe features you might have missed": A clickbait title for a blog post you don't want to miss! :wink:
Jeroen Ooms [describes five recent additions](/blog/2026/06/07/r-universe-updates/) to the R-universe platform:

- Social media cards that actually look good 
- PACKAGES.rds support (or: implementing R internals in JavaScript) 
- Fancy sort/filter bars in the WebUI 
- For the impatient: trigger a sync manually
- Making check results easier to find and share 

In other news, R-universe user Tom Palmer also wrote about five things: ["Five tips for managing your R-universe 🚀"](https://remlapmot.github.io/post/2026/runiverse-tips/).
You won't believe the fifth one. :wink:
### Yanina Joins the 2026 Sovereign Tech Fellowship

We're excited to share that our Community Manager, Yani Bellini Saibene, [has been selected as a 2026 Sovereign Tech Fellow](https://www.sovereign.tech/news/meet-the-2026-sovereign-tech-fellows). During the fellowship, she will focus on making open source more accessible through improved contribution guidance, newcomer-focused mini-hackathons, multilingual training resources, and more sustainable localization practices across communities in the R ecosystem. These efforts will build on and extend rOpenSci's work in community building, mentorship, and open science.
### _Quinceañera_: celebrating 15 years together 

In June, we held three community events to mark rOpenSci's 15th anniversary. Across all three sessions, people shared memories of their first contribution, discussed ideas for the next 15 years, and reminded us of the thing that has always made rOpenSci what it is: how genuinely welcoming it is to everyone who wants to participate. 
There's more to come :-) Keep an eye out for what we have planned for the rest of the year.

### Coworking 

Read [all about coworking](/blog/2023/06/21/coworking/)!


- Tuesday July 7nd 2026, 09:00 Americas Pacific (16:00 UTC) ["Debugging in R"](/events/coworking-2026-07/), with [Yanina Bellini Saibene](/author/yanina-bellini-saibene/) and cohost [Shannon Pileggi](/author/shannon-pileggi).
    -   Read up on debugging in R.
    -   Meet community host, Shannon Pileggi, and discuss tips and tricks for debugging in R.

- Tuesday Tuesday August 04, 09:00 Australia Western (01:00 UTC) ["Vale and Text Linting"](/events/coworking-2026-08/), with [Steffi LaZerte](/author/steffi-lazerte) and cohost [Jonathan Carroll](/author/jonathan-carroll/).
    - Read up on text linting and learn more about vale.
    - Setup a linting framework for your projects.
    - Meet co-host, Jonathan Carroll, and discuss Vale and text linting.

- Tuesday September 01, 14:00 Europe Central (12:00 UTC) ["Getting to Know SORTEE"](/events/coworking-2026-09/), with [Steffi LaZerte](/author/steffi-lazerte) and cohost [Ed Ivimey-Cook](/author/ed-ivimey-cook/).
    - Visit SORTEE (Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology).
    - Meet co-host, Ed Ivimey-Cook, and learn more about SORTEE and how you might get involved.
 
And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!


## Software :package:


```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$Package,
    description = cran_unquote(entry$Title),
    details = cran_unquote(entry$Description),
    on_cran = !is.null(entry$cran_url),
    on_bioc = !is.null(entry$`_bioc`),
    onboarding = entry$`_metadata`$review$url %||% "",
    url = entry$RemoteUrl,
    maintainer = sub(" >.*", "", entry$Maintainer)
  )
}

registry <- "https://ropensci.r-universe.dev/api/packages/" |>
  jsonlite::read_json() |>
  purrr::map_df(tidy_package)

since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 21
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" |>
    sprintf(commits[[i]]$sha) |>
    jsonlite::read_json() |>
    purrr::map_df(function(x) {
      tibble::tibble(package = x$package, url = x$url, branch = x$branch)
    })
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !package %in% "gpg",
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE, eval=FALSE}
new$onboarding[
  new$package == "hdcuremodels"
] <- "https://github.com/ropensci/software-review/issues/692"
packages <- split(new, seq(nrow(new)))
packages <- rev(packages)
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) |>
    purrr::pluck("download_url")

  if (package$package == "eDNAjoint") {
    desc_link <- "https://raw.githubusercontent.com/ropensci/eDNAjoint/26fb2d22dcd4314cdf7d170812c6247ae51f25a0/DESCRIPTION"
  }

  withr::with_tempfile("tf", {
    download.file(desc_link, tf)
    desc <<- desc::desc(file = tf)
  })
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {
    any(x$role %in% "aut") && all(x$role != "cre")
  })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {
    any(x$role %in% "rev") && all(x$role != "cre")
  })
  rev <- purrr::map_chr(rev, function(x) {
    paste(c(x$given, x$family), collapse = " ")
  })
  if (package$package == "forcis") {
    rev <- c(rev, "Scott Chamberlain")
  }
  if (package$package == "sasquatch") {
    rev <- c("Nic Crane")
  }
  maintainer <- purrr::keep(authors, function(x) {
    any(x$role %in% "cre")
  })
  maintainer <- paste(
    c(maintainer[[1]]$given, maintainer[[1]]$family),
    collapse = " "
  )
  author_string <- sprintf("developed by %s", maintainer)

  if (length(aut) > 0) {
    author_string <- paste0(
      author_string,
      sprintf(" together with %s", nice_string(aut))
    )
  }

  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package,
    package$package,
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )

  string <- gsub("doi\\:", "https://doi.org/", string)

  if (package$on_cran) {
    string <- paste0(
      string,
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string,
      sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(
      string,
      sprintf("It has been [reviewed](%s)", package$onboarding)
    )
    if (package$package == "fireexposuR") {
      rev <- c("Sherry Zhang ", "Ronny A. Hernández Mora")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  string <- sub("twitter", "social media", string)
  paste("+", string)
}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry |>
  dplyr::rowwise() |>
  dplyr::mutate(
    owner = strsplit(urltools::path(url), "/")[[1]][1],
    repo = strsplit(urltools::path(url), "/")[[1]][2]
  ) |>
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if (
    length(info) == 0 ||
      anytime::anytime(info[[1]]$published_at) < last_newsletter
  ) {
    return(NULL)
  }

  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

The writexl package has a [new maintainer](https://github.com/ropensci/writexl/pull/98#issuecomment-4191858158), Bill Denney.
NLMR is now maintained by [Jakub Nowosad](https://github.com/ropensci/NLMR/issues/116#issuecomment-4280937012).

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub(
    "Package:",
    "",
    lines[grepl("^Package", lines)][1]
  ))
  description <- stringr::str_squish(sub(
    "Title:",
    "",
    lines[grepl("^Title", lines)][1]
  ))
  if (name == "read.abares") {
    description <- "Provides simple downloading, parsing and importing of Australian
 Bureau of Agricultural and Resource Economics and Sciences (ABARES) data sources"
  }
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login

  if (name == "Ryan Zomorrodi") {
    return("[Ryan Zomorrodi](https://ryanzomorrodi.github.io/)")
  }

  if (name == "Marc Burri") {
    return("[Marc Burri](https://marcburri.github.io/)")
  }

  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) {
    url <- paste0("http://", url)
  }
  sprintf("[%s](%s)", name, url)
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else(
      "stats" %in% purrr::map_chr(issue$labels, "name"),
      " (Stats).",
      ""
    )
  )

  dplyr::rowwise(df) |>
    dplyr::mutate(
      text = sprintf(
        "    * [%s](%s), %s. Submitted by %s. %s",
        name,
        sub("http\\:", "https:", url),
        description,
        user,
        stats
      )
    )
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state,
    .limit = Inf,
    .progress = FALSE
  )

  issues <- purrr::keep(issues, \(x) {
    label %in% purrr::map_chr(x[["labels"]], "name")
  })

  purrr::map_df(issues, tidy_issue)
}

active_issues <- purrr::map_df(
  c(
    "1/editor-checks",
    "2/seeking-reviewer(s)",
    "3/reviewer(s)-assigned",
    "4/review(s)-in-awaiting-changes",
    "5/awaiting-reviewer(s)-response",
    "6/approved"
  ),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label = "6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  }

  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) |>
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf(
  "There are %s recently closed and active submissions",
  english(nrow(issues))
)
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(
    issues$label == label,
    na.rm = TRUE
  )))
  url <- glue::glue(
    'https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A%22{label}%22'
  )
  sprintf(
    "* %s at ['%s'](%s):\n\n %s",
    no,
    label,
    url,
    paste0(
      issues$text[!is.na(issues$label)][issues$label == label],
      collapse = "\n\n"
    )
  )
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(
        sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE),
        count_label
      ),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}
parse_one_post <- function(path) {
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))

  if (lines[2] == "headless: true") {
    return(NULL)
  }

  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)

  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in%
      yaml$tags &&
      !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags &&
      !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
  )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[
  !paths %in%
    c(
      "../_index.md",
      "../2021-02-03-targets/raw_data_source.md",
      "../2021-02-03-targets/README.md"
    )
]
paths <- grep("index", paths, value = TRUE)
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }

  post <- dir[which(dir[["language"]] == main_language), ]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(
      string,
      ". ",
      sub("[\\.\\!\\?]$", "", post$description),
      "."
    )
    if (url == "/blog/2025/09/18/markdown-programmatic-parsing/") {
      string <- paste(string, ".")
    }
  } else {
    string <- paste0(string, ".")
  }
  if (post$title == "How to Assess Usage of your Package") {
    string <- paste0(
      string,
      " This post was featured in the [R Weekly Highlights podcast](https://serve.podhome.fm/episodepage/r-weekly-highlights/216) hosted by Eric Nantz and Mike Thomas."
    )
  }

  other_langs <- dir[which(dir[["language"]] != main_language), ]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf(
        "<a href='%s' lang='%s'>%s (%s)</a>",
        .x[["url"]],
        .x[["language"]],
        .x[["title"]],
        .x[["language"]]
      )
    ) |>
      toString()
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  if (post$socialImg != "") {
    if (!startsWith(post$socialImg, "https")) {
      img_file <- fs::path_file(post$socialImg)
      download.file(
        sprintf("https://ropensci.org/%s", post$socialImg),
        img_file
      )
      img_file |>
        magick::image_read() |>
        magick::image_scale("400x") |>
        magick::image_write(img_file)
    } else {
      img_file <- post$socialImg
    }

    string <- paste0(
      string,
      "\n```{=html}\n",
      sprintf(
        '{{< figure class="center" src="%s" alt="%s" width="400" >}}\n\n',
        img_file,
        post$socialAlt
      ),
      "\n```\n"
    )
  }

  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(
  posts,
  ~ any(.x[["software_peer_review"]])
)]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~ any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~ any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

- [charlatan](https://docs.ropensci.org/charlatan), create fake data in R. [Issue for volunteering](https://github.com/ropensci/charlatan/issues/150).

- [hddtools](https://docs.ropensci.org/hddtools), Tools to discover hydrological data, accessing catalogues and databases from various data providers. [Issue for volunteering](https://github.com/ropensci/hddtools/issues/36).
      

### Calls for contributions

Refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.


## Package development corner

Some useful information for R package developers. :eyes:

### goodpractice's new features and behind-the-scene notes

Software Review Lead Mark Padgham and long-time community member [Athanasia Mo Mowinckel](/author/athanasia-mo-mowinckel/) have written a blog post most relevant to package developers for two reasons:

- Learn how goodpractice, that help make your package better, has improved.
- Read how Mark and Mo collaborated, including their use of LLMs in the development process.

### Dumb Ways for an Open Source Project to Die

If you're interested in open-source software projects' survivability, you'll enjoy this [write-up by Andrew Nesbitt](https://nesbitt.io/2026/05/19/dumb-ways-for-an-open-source-project-to-die.html) shared by Yanina Bellini Saibene.

### Refactoring with Jarl: unused functions and more

Hannah Frick and Maëlle Salmon wrote ["Refactoring with Jarl: a coffee chat"](https://blog.r-hub.io/2026/06/02/jarl/) on the R-hub blog.

### A strategy for recovering data on request interruption

Gábor Csárdi summarized [recent changes to the gh package](https://gaborcsardi.org/2026/gh-1-6-0-recover-from-interruption/).
Especially interesting is his strategy for [interruptions](https://gaborcsardi.org/2026/gh-1-6-0-recover-from-interruption/#Interrupts): the user starts a long query then interrupts the process...
how to not lose the data that's already been received?
The solution is to make it accessible through `rlang::last_error()`.
More details in the [post](https://gaborcsardi.org/2026/gh-1-6-0-recover-from-interruption/#Interrupts).

### curl summer of bliss

The curl project [announced](https://daniel.haxx.se/blog/2026/06/15/curl-summer-of-bliss/) that it will not accept any vulnerability report during the month of July this year.
This is both the opportunity for maintainers to take a break, and to advertise paid curl support, in which there will be no interruption of service.

### To conventionally commit or not

Sumner Evans wrote an interesting post [criticizing the conventional commits convention](https://sumnerevans.com/posts/software-engineering/stop-using-conventional-commits/) (starting commits with e.g. `fix:` for bug fixes, `feat:` for new features, etc).

### More than .gitignore

Nelson Figueroa wrote an interesting [overview](https://nelson.cloud/.gitignore-isnt-the-only-way-to-ignore-files-in-git/) of the different ways to make Git ignore some files.

### How to work with LLMs without losing your skills

Vicki Boykis wrote an insightful post ["We should be more tired than the model"](https://vickiboykis.com/2026/05/28/we-should-be-more-tired-than-the-model/) including pratical tips such as "Starting to use the agent only after I’ve spent 20 minutes on the problem" or "Discussing an agent’s proposed implementation with another person instead".

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org). This guide will help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways such as through sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so though our signup form](/news/). Until it's time for our next newsletter, you can keep in touch with us through our [website](/), [Mastodon](https://hachyderm.io/@rOpenSci), or [LinkedIn](https://www.linkedin.com/company/ropensci/). 
See you soon!
