---
title: rOpenSci News Digest, February 2023
author:
  - Maëlle Salmon
date: '2023-02-17'
slug: ropensci-news-digest-february-2023
categories:
  - blog
tags:
  - newsletter
description: R-universe improvements, coworking, champions program kick-off, targets community call resources and recording
params:
  last_newsletter: '2023-01-20'
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### R-universe improvements!

We have changed the preferred git repo name where you host your `packages.json` registry for R-universe, see our post [New preferred repo name for r-universe registries](/blog/2023/02/07/runiverse-registry-repo/).
We now recommend to use the full lowercase name of your R-universe domain as the repo name, for instance `TileDB-Inc/tiledb-inc.r-universe.dev`.

Speaking of domains and URLs, we have [improved the URL schema for the r-universe.dev front-end](/blog/2023/01/30/runiverse-permanent-urls/) to make it even easier to find and share a link to an R package, article, API, or other resource.

### Coworking

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.

* [Tuesday, Mar 7th, 9:00 Australian Western / 01:00 UTC](/events/coworking-2023-03/) "Checking data with naniar, visdat, assertr, and skimr" *Hosted by community host [Nick Tierney](/author/nicholas-tierney/) and [Steffi LaZerte](/author/steffi-lazerte/)*
    - Explore documentation, use cases and tutorials on these packages, or check your data!
    - Talk to [Nick](/author/nicholas-tierney/) and discuss how to use these packages in your workflow.
    
* [Tuesday, Apr 4th, 14:00 European Central / 12:00 UTC](/events/coworking-2023-04/) "Working with taxonomic lists in R" *Hosted by community host [Miguel Alvarez](/author/miguel-alvarez/) and [Steffi LaZerte](/author/steffi-lazerte/)*
    - Explore R packages for working with lists (such as [taxlist](https://docs.ropensci.org/taxlist/)),
    - Organize some taxonomy in your data using taxlist,
    - Talk to [Miguel](/author/miguel-alvarez/) and discuss tips for working with taxonomic lists.
    
And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

### Community Call {targets} in Action

On January 31 we enjoyed a Community Call about the [targets package](https://books.ropensci.org/targets/).

As the name of the call indicates, we had three presentations where the speakers shared their experiences putting targets into action.

Eric Scott started the panel, sharing his [attempt to make it easy to set up targets to harness the power of HPC from the comfort of an RStudio window](https://cct-datascience.quarto.pub/harnessing-hpc-power-with-targets/).  Next, Joel Nitta talked about [using targets for bioinformatics pipelines](https://joelnitta.github.io/comm-call-bioinfo-targets/), and finally Will Landau demonstrated [how to debug {targets} pipelines](https://wlandau.github.io/targets-debug/).

You can see [the video](https://vimeo.com/795053980) and access to all the material (slides, code, and shared notes with transcription and Q&A) on the [community call webpage](/commcalls/jan2023-targets/).

*What would YOU like to hear about in an rOpenSci Community Call?*, [let us know](https://github.com/ropensci-org/community-calls)

### The Champions Program is off to a great start in 2023!

We started the 2023 activities of our champions program with two kick-off meetings. 

During these meetings, [mentors and champions](/champions/) introduced themselves, shared where they are from, what their project is about, and something that excites them about the program. We also went over the logistics of the program, the potential agenda, and scheduled activities.  

We have a diverse group of people from Argentina, Brazil, Peru, Nigeria, Turkey, Australia, France, Norway, Saudi Arabia, the United States, and the United Kingdom.

We're excited to share that **three** of our champions will **develop a new package**, **four** will **prepare and submit their package to the review process**, and **three** will **become rOpenSci reviewers**. Mentors will work with their mentees throughout the process of executing their projects.


It was lovely to see the participants' enthusiasm to learn, share, execute their projects, and give back to the community!

Stay tuned for more information about our mentors, champions, and their projects!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE, eval=FALSE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

  fix_url <- function(url) {
      if (url == "http://www.blackbawks.net") {
    url <- "https://www.blackbawks.net"
  }
  if (url == "http://trang.page/") {
    url <- "https://trang.page/"
  }
  if (url == "http://www.cynkra.com") {
    url <- "https://www.cynkra.com"
  }
  if (url == "http://iecastro.netlify.com/") {
    url <- "https://iecastro.netlify.com/"
  }
  if (url == "http://rwparsons.github.io/") {
    url <- "https://rwparsons.github.io/"
  }
  if (url == "http://marceloarayasalas.weebly.com/") {
    url <- "https://marceloarayasalas.weebly.com/"
  }
  if (url == "http://jounihelske.netlify.app/") {
    url <- "https://jounihelske.netlify.app/"
  }
    url
  }
get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("https://", url)
  sprintf("[%s](%s)", name, fix_url(url))
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )

  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    twitterImg = yaml$twitterImg %||% "",
    twitterAlt = yaml$twitterAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug
    )

  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, seq(nrow(posts)))
format_post <- function(post) {
  url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(post$date),
    stringr::str_pad(lubridate::month(post$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(post$date), 2, "0", side = "left"),
    post$slug
    )
  string <- sprintf("* [%s](%s) by %s", post$title, url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  if (url == "/blog/2023/02/07/what-does-it-mean-to-maintain-a-package") {
    string <- paste0(
      string,
      " This post was discussed on the [R Weekly highlights podcast](https://rweekly.fireside.fm/111) hosted by Eric Nantz and Mike Thomas.")
  }
  
  if (post$twitterImg != "") {
    img_file <- fs::path_file(post$twitterImg)
    download.file(sprintf("https://ropensci.org/%s", post$twitterImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$twitterAlt)
    )
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, "software_peer_review")]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, "other")]
if (length(others) > 0) {
  if (length(others) != length(posts)) cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, "tech_note")]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Call for maintainers

### Call for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/) (or listening to its discussion on the [R Weekly highlights podcast](https://rweekly.fireside.fm/111) hosted by Eric Nantz and Mike Thomas)!

- **[citesdb](https://docs.ropensci.org/citesdb/)**, a high-performance database of shipment-level CITES trade data. Provides convenient access to over 40 years and 20 million records of
endangered wildlife trade data from the Convention on International Trade
in Endangered Species of Wild Fauna and Flora, stored on a local on-disk,
out-of memory 'DuckDB' database for bulk analysis. [Issue for volunteering](https://github.com/ropensci/citesdb/issues/21).


### Call for comaintainers

- rtweet, that interfaces Twitter API, is looking for a [co-maintainer](https://github.com/ropensci/rtweet/issues/763).

- Refer to our [recent blog post](/blog/2022/10/17/maintain-or-co-maintain-an-ropensci-package/#packages-looking-for-co-maintainers) to identify other packages where help is especially wished for!

## Package development corner

Some useful tips for R package developers. :eyes:

### New CRAN guideline on CITATION file

CRAN requires CITATION files to be declared as [`bibentry` items](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/bibentry.html), and not in the previously-accepted form of [`citEntry()`](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/citEntry.html).
As an example see [the dynamite CITATION file](https://github.com/ropensci/dynamite/blob/main/inst/CITATION) which refers to the R manual as well as other associated publications.

### Reference organization: the `@family` tag

In a package with more than a few help topics, reference organization is key to the user's experience.
Alphabetical order is rarely informative. :wink:
A powerful tool for organizing both the local and pkgdown reference is the [`@family` tag](https://roxygen2.r-lib.org/articles/index-crossref.html#family).

```r
#' @family datawrangling
```

- Locally it populates the "See also" section. You can tweak the title of the See also section for it not to be, say, "Other datawrangling" but rather "Data wrangling functionality", by having in `man/roxygen/meta.R`:

```r
list(
  rd_family_title = list(datawrangling = "Data wrangling functionality")
)
```

- In the pkgdown configuration for the reference index you can use [topic matching helpers](https://pkgdown.r-lib.org/reference/build_reference.html#topic-matching) such as `has_concept()` and `lacks_concepts()`.

### Gotcha: `invisible()` needs `return()`

The official man entry for `invisible()` [says that it](https://stat.ethz.ch/R-manual/R-devel/library/base/html/invisible.html):

> Return[s] a (temporarily) invisible copy of an object"

This might make it sound like `invisible()` directly implements an active **return** function, which it does not. While not strictly wrong, this statement is potentially misleading, especially followed as it is by the *Details*:

> This function can be useful when it is desired to have functions return values which can be assigned, but which do not print when they are not assigned. 

`invisible()` will only directly *return* from a function when it is wrapped in an active `return()` statement: `return(invisible())`; or when it's the very last line of the function code. 

This is all very nicely explained by [Peter Meissner on Stack Overflow](https://stackoverflow.com/a/38905341).

### A testing pattern: adding switches to your code

A recent R-hub blog post [explains a pattern for testing your package's behavior in for instance the absence of an internet connection, without turning off wifi](https://blog.r-hub.io/2023/01/23/code-switch-escape-hatch-test/).
It's based on using an early return in a helper function based on the presence of a specifically named environment variable (`MYPKG.TEST.NOINTERNET`) and on setting that environment variable locally in a test through `withr::local_envvar()`.

### Code comments tips

Another R-hub blog post [presents tips for writing as few comments as possible, while making the code as readable as possible](https://blog.r-hub.io/2023/01/26/code-comments-self-explaining-code/).

<!-- To be curated by hand -->

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Twitter account](https://twitter.com/ropensci).
