---
title: rOpenSci News Digest, January 2023
author:
  - Maëlle Salmon
date: '2023-01-20'
slug: ropensci-news-digest-january-2023
categories: []
tags:
  - newsletter
description: keywords from the content
params:
  last_newsletter: '2022-12-16'
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### {targets} in Action Community Call

Tuesday, 31 January 2023 20:00 UTC / Tuesday, 31 January 2023 15:00 EST / Wednesday, 1st February 07:00 AEDT.

The [{targets}](https://docs.ropensci.org/targets/) package is a pipeline tool for Statistics and data science in R. With {targets}, you can maintain a reproducible workflow without repeating yourself. {targets} learns how your pipeline fits together, skips costly runtime for tasks that are already up to date, runs only the necessary computation, supports implicit parallel computing, abstracts files as R objects, and shows tangible evidence that the results match the underlying code and data.

On this call Will, Eric and Joel will share their experience putting **{targets} into action**. Eric will share with us _Using {targets} with HPC_ and Joel will talk about _Using {targets} for bioinformatics pipelines_, then Will will demonstrate _Debugging {targets} pipelines_.

More info on the [event page](/commcalls/jan2023-targets/).

### Coworking

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.

- Tuesday, Feb 7th, 9:00 Americas Pacific / 17:00 UTC](/events/coworking-2023-02/) "Setting up Continuous Integration" *Hosted by community host [Hugo Gruson](/author/hugo-gruson/) and [Steffi LaZerte](/author/steffi-lazerte/)*
    - Do some reading to learn about Continuous Integration; Setup Continuous Integration on one (or more) of your projects;
    - Talk to [Hugo](/author/hugo-gruson/) and discuss how Continuous Integration can simplify your development process and how to get set up.

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

### Code of Conduct annual review and transparency report

Find [our annual review of the rOpenSci Code of Conduct, reporting process, and internal guidelines for handling reports and enforcement](/blog/2023/01/06/conduct2023/) as well as our [transparency report](/blog/2023/01/06/transparency2022/).
We thank our two independent community members [Megan Carter](https://www.esipfed.org/about/people/megan-carter) (until June 2022) and [Kara Woo](https://karawoo.com/).

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    twitterImg = yaml$twitterImg %||% "",
    twitterAlt = yaml$twitterAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug
    )

  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, seq(nrow(posts)))
format_post <- function(post) {
  url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(post$date),
    stringr::str_pad(lubridate::month(post$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(post$date), 2, "0", side = "left"),
    post$slug
    )
  string <- sprintf("* [%s](%s) by %s", post$title, url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$twitterImg != "") {
    img_file <- fs::path_file(post$twitterImg)
    download.file(sprintf("https://ropensci.org/%s", post$twitterImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$twitterAlt)
    )
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, "software_peer_review")]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, "other")]
if (length(others) > 0) {
  if (length(others) != length(posts)) cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, "tech_note")]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Use cases

```{r usecases, eval=FALSE}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
usecases <- dplyr::filter(usecases, date >= as.Date(last_newsletter))
usecases <- split(usecases, seq(nrow(usecases)))
```

`snakecase::to_sentence_case(english(length(usecases)))` use case`if (length(usecases) > 1) "s"` of our packages and resources ha`if (length(usecases) > 1) "ve" else "s"` been reported since we sent the last newsletter.

```{r usecases2, results='asis', eval=FALSE}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", sub("\\.$", "", usecase$title), usecase$url, usecase$reporter)
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)!

## Call for maintainers

### Calls for maintainers

- **[RSelenium](https://docs.ropensci.org/RSelenium), R Bindings for 'Selenium WebDriver'**. Provides a set of R bindings for the ['Selenium 2.0 WebDriver'](https://www.selenium.dev/documentation/) using the ['JsonWireProtocol'](https://github.com/SeleniumHQ/selenium/wiki/JsonWireProtocol). 'Selenium 2.0 WebDriver' allows driving a web browser natively as a user would either locally or on a remote machine using the Selenium server it marks a leap forward in terms of web browser automation. [Issue for volunteering](https://github.com/ropensci/RSelenium/issues/261).

- **[elastic](https://docs.ropensci.org/elastic), General Purpose Interface to 'Elasticsearch'**. Connect to 'Elasticsearch', a 'NoSQL' database built on the 'Java' Virtual Machine. Interacts with the ['Elasticsearch' 'HTTP' API](https://www.elastic.co/elasticsearch/), including functions for setting connection details to 'Elasticsearch' instances, loading bulk data, searching for documents with both 'HTTP' query variables and 'JSON' based body requests. [Issue for volunteering](https://github.com/ropensci/elastic/issues/292).

### Call for comaintainers

Refer to our [recent blog post](/blog/2022/10/17/maintain-or-co-maintain-an-ropensci-package/#packages-looking-for-co-maintainers) to identify packages where help is especially wished for!

## Package development corner

Some useful tips for R package developers. :eyes:

### Bad code? Good code?

Why not feel code shame when looking at older, less good code of yours?
In case you might ignore it, this 2015 blog post by David Robinson underlines how important it is to not code shame anyone lest they lose the courage to keep coding and improving: ["A Million Lines of Bad Code"](http://varianceexplained.org/programming/bad-code/).
And remember [testthat always believes in you](https://github.com/r-lib/testthat/blob/9cd6e01be008376b1f7f2d8d528d725b87c0d01e/R/praise.R#L50-L57). :wink:

### Write an R Package from R Markdown?

If you ever dreamed of [writing an R package from R Markdown](https://yihui.org/en/2023/01/litr-package/), check out the [fusen](https://thinkr-open.github.io/fusen/) package by Sébastien Rochette or the [litr](https://faculty.marshall.usc.edu/jacob-bien/litr/docs/) package by Jacob Bien and Patrick Vossler.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Twitter account](https://twitter.com/ropensci).
