---
slug: "news-july-2025"
title: rOpenSci News Digest, July 2025
author:
  - The rOpenSci Team
date: '2025-07-23'
tags:
  - newsletter
description: New champions cohort, useR! 2025
params:
  last_newsletter: '2025-06-27'
  doi: "10.59350/rgcqq-rg311"
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(trimws(...), sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### Open Science with a Latin American Identity: Meet the New Cohort of the rOpenSci Champions Program

We’re very excited to introduce the new rOpenSci [Champions](/champions)! 
This cohort will participate in the program and carry out their work in Spanish, allowing us to continue strengthening the open science and research software development community in this language. 
We’re excited about the projects they’ll be developing, which tackle real-world challenges across diverse disciplines and regions of Latin America.

Read all about the new Champions and their projects in our [blog post](/blog/2025/07/03/champions-2025/).

### rOpenSci at useR! 2025

August 1st, [useR! 2025 Virtual](https://user2025.r-project.org/program/virtual/) with **free registration**. 

August 8th through 10th, 2025 [useR! 2025](https://user2025.r-project.org) will take place in Penn Pavilion, Duke University, Durham, North Carolina. 

**rOpenSci will be there!**

* [Yanina Bellini Saibene](/author/yanina-bellini-saibene/) will be presenting her keynote ["We R Together.  How to learn, use and improve a programming language as a community"](https://user2025.r-project.org/program/in-person/keynotes/yanina-bellini-saibene) on Sat, Aug 9, 2025 - 17:00–18:00 (EDT)

* [Will Landau](/author/will-landau/) will be presenting his keynote ["Poweful simulation pipelines with {targets}"](https://user2025.r-project.org/program/in-person/keynotes/will-landau) on Sun, Aug 10, 2025 - 15:00-16:00 (EDT)

* [Noam Ross](/author/noam-ross/) will be presenting "Curating a Community of Packages: Lessons from a Decade of rOpenSci Peer Review" on Sat, Aug 9, 2025 - 13:00–14:10 (EDT)

* Our Champion [Erika Siregar](/author/erika-siregar/) is presenting "rPlaywright: Bringing Playwright’s Power to R for Scalable Web Automation", Fri, Aug 8, 2025 - 18:15–19:30 (EDT)

* Our Champion [Andrea Gomez Vargas](/author/andrea-gomez-vargas/) is presenting "ARcenso: A Package Born from Chaos, Powered by Community" on Sat, Aug 9, 2025 - 13:00–14:10 (EDT)

* [Luis D. Verde Arregoita](/author/luis-d.-verde-arregoitia/) is presenting "Bringing the fun of hex stickers to your R session" on Sun, Aug 10, 2025 - 10:30–12:00 (EDT)

We are looking forward to seeing you there!

### Coworking 

Read [all about coworking](/blog/2023/06/21/coworking/)!


- Tuesday August 5th, 2025 9:00 AM Australia Western (01:00 UTC), ["Drawing publication-ready plots using R"](/events/coworking-2025-08/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Emi Tanaka](/author/emi-tanaka/).
    - Work on some figures and plots for your own work
    - Explore ways of making publication-ready plots in R
    - Chat with  Emi Tanaka about making publication-ready plots in R
    
And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE,eval=TRUE}
packages <- split(new, seq(nrow(new)))
packages <- rev(packages)
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  
  if (package$package == "eDNAjoint") {
    desc_link <- "https://raw.githubusercontent.com/ropensci/eDNAjoint/26fb2d22dcd4314cdf7d170812c6247ae51f25a0/DESCRIPTION"
  }
  
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  if (package$package == "forcis") {
    rev <- c(rev, "Scott Chamberlain")
  }
  if (package$package == "geotargets") {
    rev <- c("Anthony Martinez", "Denisse Fierro Arcos")
  }
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )

  string <- gsub("doi\\:", "https://doi.org/", string)
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (package$package == "fireexposuR") {
      rev <- c("Sherry Zhang ", "Ronny A. Hernández Mora")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  string <- sub("twitter", "social media", string)
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  if (name == "read.abares") {
    description <- "Provides simple downloading, parsing and importing of Australian
 Bureau of Agricultural and Resource Economics and Sciences (ABARES) data sources"
  }
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login

  if (name == "Ryan Zomorrodi") {
    return("[Ryan Zomorrodi](https://ryanzomorrodi.github.io/)")
  }

  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state,
    .limit = Inf,
    .progress = FALSE
  )

  issues <- purrr::keep(issues, \(x) label %in% purrr::map_chr(x[["labels"]], "name"))
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- glue::glue('https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A%22{label}%22')
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))

  if (lines[2] == "headless: true") {
    return(NULL)
  }

  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md",  "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("[\\.\\!\\?]$", "", post$description), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$socialImg != "") {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string, "\n```{=html}\n",
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt),
      "\n```\n"
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).


* [USAboundaries](https://docs.ropensci.org/USAboundaries/) (and USAboundariesdata), historical and contemporary boundaries of the United States of America. [Issue for volunteering](https://github.com/ropensci/USAboundaries/issues/50).

### Calls for contributions

Refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.


## Package development corner

Some useful tips for R package developers. :eyes:

### Feedback wanted! CRAN Task View proposal for "Package Development and Maintenance"

Sharing a request by Heather Turner and Lluís Revilla.

There is a new CRAN Task View [proposal](https://github.com/hturner/pkg-dev-ctv/blob/main/proposal.md) for "Package Development and Maintenance". 
This task view lists and comments on packages useful for package development and maintenance. The authors and the CRAN task View editorial team are looking for feedback.
In particular, they'd especially like your feedback if you're new to package development or find the idea of creating and maintaining a package intimidating (to avoid the ["curse of knowledge"](https://en.wikipedia.org/wiki/Curse_of_knowledge)).

Regarding this task view, they'd like to know:

- Does it have most of the information you need to accomplish all the steps through development until your package is released? 
- Does this help you understand which tools and packages can help you maintain your package working ?
- Is there any topic or step we missed? Or after reading the relevant topic section you still don't feel confident on how to do it?

Read the [proposal](https://github.com/hturner/pkg-dev-ctv/blob/main/proposal.md).
Please submit your feedback directly in [the repository](https://github.com/hturner/pkg-dev-ctv/issues?q=sort%3Aupdated-desc+is%3Aissue+is%3Aopen) or send the feedback through private channels to [Lluís Revilla](https://llrs.dev) (email, Slack, ...)

### Cache values with automatic pruning: the cachem package

If you're writing a wrapper for an API that has no regular caching headers, 
you might be interested in the [cachem package](https://github.com/r-lib/cachem/) for caching.
Thanks to Tan Ho and Pieter Huybrechts for the tip.
See its use in [comtradr](https://github.com/ropensci/comtradr/blob/f5b2715bf6dddcaff039c3613e8f4ecb0607e5c0/R/zzz.R#L58) by Paul Bochtler.

### A debugging tip

Miles McBain shared an interesting debugging tip in his post [Dive()ing into the hunt #rstats](https://milesmcbain.micro.blog/2025/07/04/diveing-into-the-hunt-rstats.html), using a custom function to implement something like a 'row-wise data debug' mode!

### How to prepare examples in your package for a CRAN submission

The CRAN cookbook by Jasmine Daly and Beni Altmann  features a section called ["Structuring of Examples"](https://contributor.r-project.org/cran-cookbook/general_issues.html#structuring-of-examples) mentioning dontrun and friends and their use cases.
Thanks to Hugo Gruson for the find!

### Don't ask "any updates on this?" on GitHub

Carlos Scheidegger, a developer of Quarto, wrote an interesting [Bluesky thread](https://bsky.app/profile/cscheid.net/post/3ls25wg6mnk2y) about why it's not considerate to write "any update on this?" on open bug reports or feature requests.

Carlos suggests that better alternatives are to "add _new_ information to the post about how you want to use it, or other workarounds" which might give developers ideas for different fixes, or to simply upvote the request. But avoid "any updates?", which is hard on the developer because all they can say is "no", and they know that will disappoint people. 

### LLLM helpers: the chores package

The [chores package](https://simonpcouch.github.io/chores/index.html) by Simon Couch provides LLM helpers for tasks that are hard to automate, 
like converting your messages to using the cli R package.
Its thoughtful documentation includes guidance on [creating your own helpers](https://simonpcouch.github.io/chores/articles/custom.html).

### LLM tools for R

Speaking of LLM tooling in R, Luis D. Verde Arregoitia  maintains an [extensive guide](https://luisdva.github.io/llmsr-book/).



## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).