---
title: 'rOpenSci News Digest, July 2023'
author: 'The rOpenSci Team'
date: '2023-07-21'
tags:
  - newsletter
slug: news-july-2023
description: CRAN packages indexed on R-universe, community call, coworking
params:
  last_newsletter: "2023-06-22"
---


```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### CRAN packages on R-universe

#### All CRAN packages indexed on R-universe!

We have now indexed all CRAN packages on r-universe.dev. You can use the r-universe [powersearch](/blog/2023/02/27/runiverse-discovering/#level-1-searching-the-entire-r-ecosystem) to find any CRAN package with a particular topic, author, datasets etc.

#### R-universe cran.dev shortlinks

You can now use cran.dev shortlinks: `https://cran.dev/{package}`, `https://docs.cran.dev/{package}` to fly directly to the r-universe dev page or html manual for any CRAN package!

Example: <https://cran.dev/xml2> and <https://docs.cran.dev/xml2>.

### Community call: Mentoring & training program for Scientific Open Source Champions

Tuesday, 25 July 2023 14:00 UTC. [More info](/commcalls/july2023-championprogram/).

Champions programs are designed to identify, recognize, and reward emerging leaders within a community. The [rOpenSci Champions Program](/champions/) is part of a series of activities and projects we are carrying out to ensure our research software serves everyone in our communities, which means that it needs to be sustainable and open, and built by and for all groups.

On this [call](/commcalls/july2023-championprogram/) _Beatriz Milz_, _Victor Ordu_ and _Carolina Pradier_ will share their experience of being rOpenSci mentors and champions. 
We will highlight the benefits of being part of the program for you and for your community, what kind of learning, 
activities and opportunities an open source community champions program provides. 
_Yani_ will present the details of our Champion Program and answer all your question about it.


### Coworking

Read [all about coworking](/blog/2023/06/21/coworking/) in our new [post](/blog/2023/06/21/coworking/)!

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.

- Tuesday, August 1st, 09:00 Americas Pacific (16:00 UTC), ["Spatial data in R"](/events/coworking-2023-08/) *Hosted by [Mike Mahoney](/author/mike-mahoney) and [Steffi LaZerte](/author/steffi-lazerte/)*
    - Work on a project related to spatial data.
    - Explore R packages for working with spatial data.

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!



## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE, eval=FALSE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
packages <- packages[purrr::map_chr(packages, "package") != "commonmark"]
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$socialImg != "") {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt)
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  if (length(others) != length(posts)) cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Use cases

```{r usecases}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
usecases <- dplyr::filter(usecases, date >= as.Date(last_newsletter))
usecases <- split(usecases, seq(nrow(usecases)))
```

`r snakecase::to_sentence_case(english(length(usecases)))` use case`r if (length(usecases) > 1) "s"` of our packages and resources ha`r if (length(usecases) > 1) "ve" else "s"` been reported since we sent the last newsletter.

```{r usecases2, results='asis'}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", sub("\\.$", "", usecase$title), usecase$url, usecase$reporter)
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)!

## Call for maintainers

<!--IF CALL
* [our guidance on _Changing package maintainers_](https://devguide.ropensci.org/changing-maintainers.html)
* [our _Package Curation Policy_](https://devguide.ropensci.org/curationpolicy.html)

IF NO CALL
There are no open calls for new maintainers at this point but you can refer to our [contributing guide](https://contributing.ropensci.org/) for finding ways to get involved!
As the maintainer of an rOpenSci package, feel free to contact us on Slack or email `info@ropensci.org` to get your call for maintainer featured in the next newsletter. -->

## Package development corner

Some useful tips for R package developers. :eyes:

### Check for partial matching as part of your testing & CI process

You can change your options to generate warnings on partial matching:

```r
options(
  warnPartialMatchAttr = TRUE,
  warnPartialMatchDollar = TRUE,
  warnPartialMatchArgs = TRUE 
)
```

But your options and [Rprofile](https://rstats.wtf/r-startup.html#rprofile) are ignored by R CMD check.

A way to still get warnings in your continuous integration is to integrate it to testthat by creating a [special `setup-options.R` file](https://blog.r-hub.io/2020/11/18/testthat-utility-belt/), as in https://github.com/epiverse-trace/packagetemplate/pull/61. This file could also be used to change any other options (but use with caution to not create situations where it's too difficult to debug).

Thanks to [Hugo Gruson](https://hugogruson.fr/) for this tip!

### Change in R-devel for comparing versions

Does your package compare version with code snippets like `some_version >= 1.5`?
If so, you might encounter the warning below with R-devel.

```
Condition
  Warning in `.make_numeric_version()`:
  invalid non-character version specification 'x' (type: double)
```

[Reference](https://bugs.r-project.org/show_bug.cgi?id=18548), [toot](https://mastodon.social/@zkamvar@fosstodon.org/110634968544104625).

> To avoid it, you need to make your comparison a character vector:
`some_version >= '1.5'` 

[Example commit fixing this, in usethis](https://github.com/r-lib/usethis/commit/a5e80b516154309dd28d09e642ea7e7b6c7e723a).

Thanks to [Zhian Kamvar](https://zkamvar.netlify.app/) for this!

### How to take vacation as an open-source maintainer

Some inspiration by Will Landau: https://mastodon.social/@landau@fosstodon.org/110636695153363955
Take care, y'all.

<!-- To be curated by hand -->

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
