---
slug: "news-february-2026"
title: rOpenSci News Digest, February 2026
author:
  - The rOpenSci Team
date: '2026-02-17'
tags:
  - newsletter
description: coworking, new package, package news
params:
  last_newsletter: '2026-01-29'
  doi: "10.59350/fe3c2-6zc80"
---

```{r setup, include=FALSE}
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(trimws(...), sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ
### Strong Engagement in the Champions Program Call

The Champions Program call closed on February 23, and the response _was fantastic_. 
We received **81 Champion** and **14 Mentor applications** from **23 countries**, 
with **74%** of applicants proposing to develop a **new package**. 
We’re now kicking off the selection process, starting with mentors so they can support the evaluation of Champion proposals. 
Confirmation emails have already been sent to all applicants. 
Thank you to everyone who applied!



### Coworking 

Read [all about coworking](/blog/2023/06/21/coworking/)!

-  Tuesday March 3rd 2026, 14:00 Europe Central (13:00 UTC), ["rOpenSci Editors"](/events/coworking-2026-03/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Mark Padgham](/author/mark-padgham/).  
    -   Learn more about [rOpenSci Software Peer Review](/software-review/)
    -   Meet cohost Mark Padgham, and discuss what it’s like being an Editor for rOpenSci Software Peer Review.
- Tuesday April 7th 2026, 9:00 Americas Pacific (16:00 UTC) ["Getting to know the CSID Network"](/events/coworking-2026-04/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Irene Ramos](/author/irene-ramos/).  
    -   Learn more about the [CSID Network](https://csidnet.org/)
    -   Meet cohost, Irene Ramos, and learn more about the CSID Network and how you might get involved.
- Tuesday May 5th 2026, 9:00 Australia Western (01:00 UTC) ["Code Review with rOpenSci"](/events/coworking-2026-05/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Liz Hare](/author/liz-hare/).  
    -   Explore resources for Code Review
    -   Sign up to volunteer to do [software peer-review](https://airtable.com/app8dssb6a7PG6Vwj/shrnfDI2S9uuyxtDw) at rOpenSci
    -   Meet cohost, Liz Hare, and discuss resources for Code Review with rOpenSci.

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

### Community Dashboard for your own R-Universe

Last year we published a [blog post](/blog/2025/09/11/repometrics-call-help) on an [organization-level dashboard](https://ropensci-review-tools.github.io/repometrics-demo/) for all rOpenSci packages and community contributors.
We have now developed the [orgmetricsDashboard repository](https://github.com/ropensci-review-tools/orgmetricsDashboard) to enable anybody to deploy their own organizational dashboard directly from their [R-Universe repository](https://docs.r-universe.dev/publish/set-up.html).
The only input needed is an [R-Universe "packages.json" file](https://docs.r-universe.dev/publish/set-up.html#registry-file), and you can deploy either via GitHub actions, or a local Docker workflow.
Try it out!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$Package,
    description = cran_unquote(entry$Title),
    details = cran_unquote(entry$Description),
    on_cran = !is.null(entry$cran_url),
    on_bioc = !is.null(entry$`_bioc`),
    onboarding = entry$`_metadata`$review$url %||% "",
    url = entry$RemoteUrl,
    maintainer = sub(" >.*", "", entry$Maintainer)
    
  )
}

registry <- "https://ropensci.r-universe.dev/api/packages/" |>
  jsonlite::read_json() |>
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" |>
    sprintf(commits[[i]]$sha) |>
    jsonlite::read_json() |>
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !package %in% "gpg",
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE,eval=TRUE}
new$onboarding[new$package == "hdcuremodels"] <- "https://github.com/ropensci/software-review/issues/692"
packages <- split(new, seq(nrow(new)))
packages <- rev(packages)
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) |>
    purrr::pluck("download_url")
  
  if (package$package == "eDNAjoint") {
    desc_link <- "https://raw.githubusercontent.com/ropensci/eDNAjoint/26fb2d22dcd4314cdf7d170812c6247ae51f25a0/DESCRIPTION"
  }
  
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(c(x$given, x$family), collapse = " "))
  if (package$package == "forcis") {
    rev <- c(rev, "Scott Chamberlain")
  }
  if (package$package == "sasquatch") {
    rev <- c("Nic Crane")
  }
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )

  string <- gsub("doi\\:", "https://doi.org/", string)
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (package$package == "fireexposuR") {
      rev <- c("Sherry Zhang ", "Ronny A. Hernández Mora")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  string <- sub("twitter", "social media", string)
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry |>
  dplyr::rowwise() |>
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) |>
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

Post on dfms release: [Releasing dfms 1.0: Fast and Feature-Rich Estimation of Dynamic Factor Models in R](https://sebkrantz.github.io/Rblog/2026/01/29/releasing-dfms-1-0-fast-and-feature-rich-estimation-of-dynamic-factor-models-in-r/).

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  if (name == "read.abares") {
    description <- "Provides simple downloading, parsing and importing of Australian
 Bureau of Agricultural and Resource Economics and Sciences (ABARES) data sources"
  }
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login

  if (name == "Ryan Zomorrodi") {
    return("[Ryan Zomorrodi](https://ryanzomorrodi.github.io/)")
  }

  if (name == "Marc Burri") {
    return("[Marc Burri](https://marcburri.github.io/)")
  }

  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) |>
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, sub("http\\:", "https:", url), description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state,
    .limit = Inf,
    .progress = FALSE
  )

  issues <- purrr::keep(issues, \(x) label %in% purrr::map_chr(x[["labels"]], "name"))
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) |>
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- glue::glue('https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A%22{label}%22')
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))

  if (lines[2] == "headless: true") {
    return(NULL)
  }

  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md",  "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
paths <- grep("index", paths, value = TRUE)
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("[\\.\\!\\?]$", "", post$description), ".")
        if (url == "/blog/2025/09/18/markdown-programmatic-parsing/") {
      string <- paste(string, ".")
    }
  } else {
    string <- paste0(string, ".")  
  }
    if (post$title == "How to Assess Usage of your Package") {
    string <- paste0(string, " This post was featured in the [R Weekly Highlights podcast](https://serve.podhome.fm/episodepage/r-weekly-highlights/216) hosted by Eric Nantz and Mike Thomas.")
  }
  if (post$socialImg != "") {
    if (!startsWith(post$socialImg, "https")) {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file |> magick::image_read() |> magick::image_scale("400x") |> magick::image_write(img_file)
    } else {
      img_file <- post$socialImg
    }

    string <- paste0(
      string, "\n```{=html}\n",
      sprintf('{{< figure class="center" src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt),
      "\n```\n"
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) |> 
      toString()
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

- [NLMR](https://docs.ropensci.org/NLMR), R package to simulate neutral landscape models. [Issue for volunteering](https://github.com/ropensci/NLMR/issues/116).

- [landscapetools](https://docs.ropensci.org/landscapetools), R package for some of the less-glamorous tasks involved in landscape analysis. [Issue for volunteering](https://github.com/ropensci/landscapetools/issues/48).

- [hddtools](https://docs.ropensci.org/hddtools), Tools to discover hydrological data, accessing catalogues and databases from various data providers. [Issue for volunteering](https://github.com/ropensci/hddtools/issues/36).
        
- [qualtRics](https://docs.ropensci.org/qualtRics/), download Qualtrics survey data. [Issue for volunteering](https://github.com/ropensci/qualtRics/issues/383).

- [photosearcher](https://docs.ropensci.org/photosearcher), searches Flickr for photographs and metadata. [Issue for volunteering](https://github.com/ropensci/photosearcher/issues/22).


### Calls for contributions

Refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.


## Package development corner

Some useful tips for R package developers. :eyes:

### The R Foundation on OSS

The R Foundation answered a [Call for Evidence from the EU about open-source](https://ec.europa.eu/info/law/better-regulation/have-your-say/initiatives/16213-European-Open-Digital-Ecosystems_en), 
after asking for examples of the added value of R in the public or private sectors.
You can read the [full response of the R Foundation](https://ec.europa.eu/info/law/better-regulation/have-your-say/initiatives/16213-European-Open-Digital-Ecosystems/F33370610_en), including a discussion of useful ideas for the future of R and open-source in general.

### On sitrep (situation report) functions

Athanasia Monika Mowinckel wrote a post on [sitrep functions](https://drmowinckels.io/blog/2026/sitrep-functions/):
functions that help the user check their setup and provide them with useful diagnostics.
The blog post includes two examples.

### New version of the jarl CLI

The jarl CLI by Etienne Bacher "finds inefficient, hard-to-read, and suspicious patterns of R code across dozens of files and thousands of lines of code in milliseconds".
Read about the new features in [jarl 0.4.0](https://www.etiennebacher.com/posts/2026-02-03-jarl-0.4.0/), like the ability to find "unreachable code".

### futurize R package

Henrik Bengtsson released [futurize](https://www.jottr.org/2026/01/22/futurize-0.1.0/) that allows you to parallelize execution with minimal changes to code, by simply adding a call to `futurize()`: `y <- lapply(x, fcn) |> futurize()`.

### On AI agents and open-source

Scott Shambaugh, an open-source maintainer, wrote about his being targeted by an AI agent after closing a PR for an issue that was intended for human newbies.
Initial blog post: ["An AI Agent Published a Hit Piece on Me"](https://theshamblog.com/an-ai-agent-published-a-hit-piece-on-me/), [Updated](https://theshamblog.com/an-ai-agent-published-a-hit-piece-on-me-part-2/).

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
