---
title: rOpenSci News Digest, October 2021
author:
  - The rOpenSci Team
date: '2021-10-22'
slug: ropensci-news-digest-october-2021
categories: []
tags: 
  - newsletter
description: Community call about Statistical Package Standards at rOpenSci, co-working with pkgcheck intro, NHS-R, 3 new editors, new packages and package news
output:
  html_document:
    keep_md: yes
params:
  last_newsletter: '2021-09-30'
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### Community call about the statistical software review project

A fantastic community call is coming up on Tuesday, 26 October 2021 18:00 UTC: [**Expanding Software Peer Review: Statistical Package Standards at rOpenSci**](/commcalls/oct2021-statsreview02/)!

This 1-hour community call will address the bigger picture of **how our community-informed development of standards for statistical packages meets a critical need of stakeholders**. 
[**Noam Ross**](/author/noam-ross/) (EcoHealth Alliance and rOpenSci Software Review Lead) will catch everyone up on the project. 
[**Rebecca Killick**](/author/rebecca-killick/) (Lancaster University and rOpenSci Statistical Software Peer Review advisory committee) will offer insights into standardisation and the potential role our program might play in the future of statistical software and open source software in general. 
[**Juliane Manitz**](/author/juliane-manitz/) (EMD Serono and R Validation Hub) will offer a perspective on the use of open source software in regulated environments. 
[**Christoph Sax**](/author/christoph-sax/) (cynkra) will share his experience as the first person to submit a package, tsbox, for review and aligning his software with our standards.

Follow the [community call page link](/commcalls/oct2021-statsreview02/) for practical information. 
**Everyone is welcome, no RSVP needed!**

### November 2 social coworking and office hours

Join us for [monthly coworking Tuesday, November 2nd at 16:00 UTC](/events/coworking-2021-11/)! No RSVP needed. In addition to the usual quiet and noisy rooms (for Q&A and meeting peers), [Mark Pagham](/author/mark-padgham/), co-Lead of our [Statistical Software Peer Review project](/stat-software-review/), will be there to help folks check their own package with our new automated checking system.

### rOpenSci at NHS-R conference

Jeroen Ooms, rOpenSci Lead Infrastructure Engineer, will give a talk on [The r-universe project](/r-universe/) at the [NHS-R 2021 virtual conference](https://nhsrcommunity.com/nhs-r-community-conference-2021/) on November 9. [Details](/events/nhsr-jeroen/).

Will Landau, community member, will give a workshop on Reproducible computation at scale in R with [targets](https://docs.ropensci.org/targets/) on [November 4](https://nhsrcommunity.com/nhs-r-conference-2021-workshop-week/).

The conference is free with registration. [Conference website](https://nhsrcommunity.com/nhs-r-community-conference-2021/).

Find out about more [events](/events). You can subscribe to rOpenSci-lead events via our [ICS calendar](/events/index.ics) (copy the link, add it to your calendar service e.g. [Thunderbird](https://support.mozilla.org/en-US/kb/creating-new-calendars#w_on-the-network-connect-to-your-online-calendars) or Google Calendar). Please note that if you use Google Calendar you have to [check synchronization is turned on](https://calendar.google.com/calendar/u/0/syncselect); furthermore, synchronization might be delayed.

### Three new editors for software review

We are excited to welcome Emily Riederer, Adam Sparks, and Jeff Hollister to our team of [Associate Editors](/software-review/#editors) for rOpenSci [Software Peer Review](/software-review/). 
They join Laura DeCicco, Julia Gustavsen, Anna Krystalli, Mauro Lepore, Karthik Ram, Noam Ross, Maëlle Salmon, and Melina Vidoni.
Find out more about the three new editors in our [introduction blog post](/blog/2021/10/12/editors2021/).

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
new <- new[2,]
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", toString(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", toString(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- toString(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the latest newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = toString(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    twitterImg = yaml$twitterImg %||% "",
    twitterAlt = yaml$twitterAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug
    )

  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, seq(nrow(posts)))
format_post <- function(post) {
  url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(post$date),
    stringr::str_pad(lubridate::month(post$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(post$date), 2, "0", side = "left"),
    post$slug
    )
  string <- sprintf("* [%s](%s) by %s", post$title, url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$twitterImg != "") {
    img_file <- fs::path_file(post$twitterImg)
    download.file(sprintf("https://ropensci.org/%s", post$twitterImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$twitterAlt)
    )
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, "software_peer_review")]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, "other")]
if (length(others) > 0) {
  if (length(others) != length(posts)) cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, "tech_note")]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Use cases

```{r usecases}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
usecases <- dplyr::filter(usecases, date >= as.Date(last_newsletter))
usecases <- split(usecases, seq(nrow(usecases)))
```

`r snakecase::to_sentence_case(english(length(usecases)))` use cases of our packages and resources have been reported since we sent the last newsletter.

```{r usecases2, results='asis'}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", sub("\\.$", "", usecase$title), usecase$url, usecase$reporter)
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)!

## Call for maintainers

There's no open call for new maintainers at this point but you can refer to our [contributing guide](https://contributing.ropensci.org/) for finding ways to get involved!
As the maintainer of an rOpenSci package, feel free to contact us on Slack or email `info@ropensci.org` to get your call for maintainer featured in the next newsletter. 

## Package development corner

Some useful tips for R package developers. :eyes:

### Naming your package

Do you have a fantastic idea and plan for a package, but no name for it yet?
We have some tips on this topic in our [dev guide](https://devguide.ropensci.org/building.html#naming-your-package)!

* We strongly recommend short, descriptive names in lower case. If your package deals with one or more commercial services, please make sure the name does not violate branding guidelines. You can check if your package name is available, informative and not offensive by using the [`available` package](https://github.com/ropenscilabs/available). In particular, do _not_ choose a package name that's already used on CRAN or Bioconductor.

* A more unique package name might be easier to track (for you and us to assess package use) and search (for users to find it and to find answers to their questions). Obviously a _too_ unique package name might make the package less discoverable (e.g. it might be an argument for naming your package [geojson](https://github.com/ropensci/geojson)).
Also note that removing random letter e.g. vowels from a common word to create a package name might make it less easy to remember!

* Find other interesting aspects of naming your package [in this blog post by Nick Tierney](https://www.njtierney.com/post/2018/06/20/naming-things/), and in case you change your mind, find out [how to rename your package in this other blog post of Nick's](https://www.njtierney.com/post/2017/10/27/change-pkg-name/).

### Two testthat tips

The testthat package has been supporting [snapshot tests](https://www.tidyverse.org/blog/2020/10/testthat-3-0-0/#snapshot-testing) that "record expected output in a separate human-readable file instead of using code to describe what the expected output looks like." since its version 3.0.0.
Their implementation is now stable.
Note that they are skipped by default on CRAN.

Now what about _input files_ for tests? 
Maybe you can [create fake ones on the fly (and delete them)](https://blog.r-hub.io/2020/11/18/testthat-utility-belt/#create-fake-folders-and-text-files-from-your-tests).
Or you can store them under `tests/testthat` and use [`testthat::test_path()`](https://testthat.r-lib.org/reference/test_path.html) when using them, to get a file path that "both interactively and during tests".
So to use `tests/testthat/examples/thing` you'd write `testthat::test_path("examples", "thing")`.
Note that the vcr package for HTTP testing has a similar function you can use to locate cassettes, [vcr_test_path()](https://docs.ropensci.org/vcr/reference/vcr_test_path.html).

### Passing CRAN URL checks

CRAN [checks URL validity](https://blog.r-hub.io/2020/12/01/url-checks/).
As recently [reminded by Jenny Bryan on Twitter](https://twitter.com/JennyBryan/status/1450186196395520005), the [urlchecker package by Jim Hester](https://github.com/r-lib/urlchecker/) is a great tool to help your package pass these checks.

### Styling your package

How to enforce a style guide for your package, without too much human effort?
Two suggestions:

* Having a [GitHub Action workflow that allows you to call styler](https://github.com/r-lib/actions/tree/master/examples#commands-workflow) via typing `/style` in a PR comment (or amend the logic to have it work via applying a label to the PR). Or apply this idea to another continuous integration service! If for any reason your style differs from styler's default, follow one of [styler articles for developers](https://styler.r-lib.org/dev/articles/) and tweak the workflow. [Example](https://github.com/ropensci-review-tools/pkgcheck/blob/65fa9d4e77837bd122171d46177cb7db8028ac2e/.github/workflows/pr-commands.yaml#L60-L64), corresponding [style guide package](https://github.com/ropensci-review-tools/spaceout).

* Using a service like [CodeFactor](https://www.codefactor.io) that integrates with GitHub directly. Thanks to [Rich FitzJohn](/author/rich-fitzjohn/) for suggesting this in [rOpenSci semi-open Slack workspace](https://contributing.ropensci.org/resources.html#channels).

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Twitter account](https://twitter.com/ropensci).
