---
title: How to Translate a Hugo Blog Post with Babeldown
author: Maëlle Salmon
date: '2023-09-05'
slug: how-to-translate-a-hugo-blog-post-with-babeldown
categories: []
tags:
  - tech notes
  - multilingual
---

As part of our [multilingual publishing project](/multilingual-publishing/), and with [funding from the R Consortium](https://www.r-consortium.org/all-projects/awarded-projects/2022-group-2), we've worked on the R package [babeldown](https://docs.ropensci.org/babeldown/) for translation of Markdown based content through DeepL API.
In this tech note, we'll see how to use babeldown to translate a Hugo blog post!


## Motivation

Translating a blog post from your R console is not only more comfortable when you've already written said blog post in R.
With babeldown, compared to copy-pasting the content of a blog post into some translation service, you should not break the Markdown syntax, and code chunks won't be translated.
Under the hood, babeldown uses [tinkr](https://docs.ropensci.org/) to produce XML to send to DeepL API, flagging some tags as not to be translated, and to convert the XML translated by DeepL into Markdown again.

Now, obviously, machine-translated content isn't perfect yet!
A human or two will need to review and amend the translation.
Why not have the humans translate the post from scratch then?
We have observed that editing an automatic translation is faster than translating the whole post, and that it frees up mental space for implementing translation rules such as gender-neutral phrasing.

## Setup

### Pre-requisites on the Hugo website

`babeldown::deepl_translate_hugo()` assumes the Hugo website uses

- leaf bundles (each post in a folder, `content/path-to-leaf-bundle/index.md`);
- multilingualism so that a post in say Spanish lives in `content/path-to-leaf-bundle/index.es.md`.

We could envision extending this to other Hugo multilingual setups, please open an issue in [babeldown repository](https://github.com/ropensci-review-tools/babeldown/) if you'd be interested in using babeldown for a Hugo website you set up differently.

### DeepL pre-requisites

First check your desired source and target languages are supported by DeepL API!
The API docs has a list of [target languages](https://www.deepl.com/docs-api/general/get-languages).
Finding a list of _source languages_ might require going through the [Help Center](https://support.deepl.com/).
Or it might just be quicker to register for an account and use `babeldown::deepl_languages()`.
We'll put the current lists below using our API key.

```{r, echo=FALSE}
Sys.setenv("DEEPL_API_URL" = "https://api.deepl.com")
Sys.setenv(DEEPL_API_KEY = keyring::key_get("deepl"))
```

```{r}
babeldown::deepl_languages(type = "target") |> knitr::kable()
```
```{r}
babeldown::deepl_languages(type = "source") |> knitr::kable()
```

Once you know you'll be able to take advantage of the DeepL API, you'll need to create an account for the [API of the translation service DeepL](https://www.deepl.com/en/docs-api/).
Even getting a free account demands registering a payment method with them.

### R pre-requisites

You'll need to install babeldown from rOpenSci R-universe:

```r
install.packages('babeldown', repos = c('https://ropensci.r-universe.dev', 'https://cloud.r-project.org'))
```

Then, in each R session, set your DeepL API key via the environment variable DEEPL_API_KEY. You could store it once and for all with the [keyring](https://r-lib.github.io/keyring/index.html) package and retrieve it in your scripts like so:

```r
Sys.setenv(DEEPL_API_KEY = keyring::key_get("deepl"))
```

Lastly, the DeepL API URL depends on your API plan. 
babeldown uses the DeepL free API URL by default. 
If you use a Pro plan, set the API URL in each R session/script via

```r
Sys.setenv("DEEPL_API_URL" = "https://api.deepl.com")
```

## Translation!

You could simply run the code below

```r
babeldown::deepl_translate_hugo(
  post_path = <path-to-post>,
  source_lang = "EN",
  target_lang = "ES",
  formality = "less" # that's we roll here!
)
```

but we'd recommend a tad more work for your own good.

### A workflow for Git and GitHub

If you use version control, having the translation as a diff is very handy!

#### In words

- Save your original blog post under the target blog post name and commit it.
- Create a new branch.
- Run `babeldown::deepl_translate_hugo()` with `force = TRUE`.
- Commit the result and open a PR. The diff for the target blog post will be the diff between the source and target language! If you have the good habit to start a new line after each sentence / sentence part, it's even better.

#### In code

With code using fs and gert (but you do you!), assuming your current directory is the root of the website folder, and also the root of the git repository, and a translation from Spanish to English...

- Save your original blog post under the target blog post name and commit it.

```r
fs::file_copy(
  file.path("content", "blog", "2023-09-15-r-universe-stars-5", "index.es.md"),
  file.path("content", "blog", "2023-09-15-r-universe-stars-5", "index.md")
)
gert::git_add(file.path("content", "blog", "2023-09-15-r-universe-stars-5", "index.md"))
gert::git_commit("Add translation placeholder")
```

- Create a new branch.

```r
gert::git_branch_create("translation-tech-note")
```

- Run `babeldown::deepl_translate_hugo()` with `force = TRUE`.

```r
babeldown::deepl_translate_hugo(
  post_path = file.path("content", "blog", "2023-09-15-r-universe-stars-5", "index.md"),
  force = TRUE,
  yaml_fields = c("title", "description", "tags"),
  source_lang = "ES",
  target_lang = "EN-US",
  formality = "less"
)
```

- Commit the result and open a PR.

```r
gert::git_add(file.path("content", "blog", "2023-09-15-r-universe-stars-5", "index.md"))
gert::git_commit("Add translation")
gert::git_push()
# and on to GitHub interface, or use `usethis::pr_init()` + `usethis::pr_push()`
```

The human translators can then a open a _second_ PR to the translation branch with their edits!

### YAML fields

By default babeldown translates the YAML fields "title" and "description". 
If you have text in more of them, use the `yaml_fields` argument of `babeldown::deepl_translate_hugo()`.

Note that if babeldown translates the title, it updates the slug.

### Glossary

Image you have a few preferences for some words -- something you'll build up over time.

```{r}
readr::read_csv(
  system.file("example-es-en.csv", package = "babeldown"), 
  show_col_types = FALSE
)
```

You can record this in a glossary in your DeepL account

```r
deepl_upsert_glossary(
  <path-to-csv-file>,
  glossary_name = "rstats-glosario",
  target_lang = "Spanish",
  source_lang = "English"
)
```

You'd use the exact same code to _update_ the glossary hence the name "upsert" for the function.
You need one glossary per source language / target language pair: the English-Spanish glossary can't be used for Spanish to English for instance.

In your `babeldown::deepl_translate_hugo()` call you then use the glossary name (here "rstats-glosario") for the `glossary` argument. 

## Conclusion

In this post we explained how to translate a Hugo blog post using babeldown.
Although the gist is to use one call to `babeldown::deepl_translate_hugo()`,
- one needs to indicate the API URL and key, 
- one can improve results by using the function's different arguments,
- we recommend pairing the translation with a Git + GitHub (or GitLab, gitea...) workflow.

babeldown has [functions](https://docs.ropensci.org/babeldown/reference/index.html) for translating Quarto book chapters, any Markdown file, any Markdown string, with similar arguments and recommended usage, so explore its reference!

We'd be happy to hear about your [use cases](/usecases/).


