---
slug: ropensci-news
title: Latest rOpenSci News Digest
author:
  - The rOpenSci Team
date: '2021-03-15'
tags:
  - newsletter
description: R-universe page, Statistical software review comm call, new packages and package news.
output:
  html_document:
    keep_md: yes
params:
  last_newsletter: "2021-02-08"
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}
```

Dear rOpenSci friends, it's time for our monthly news roundup!
Some housekeeping first:
We've been changing the infrastructure of our newsletter a bit so please update your RSS and JSON feeds to `https://ropensci.org/tags/newsletter/index.xml` and `https://ropensci.org/tags/newsletter/index.json`, respectively.
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

Our R-universe project now has [its own page on our website](/r-universe/)! :rocket:

Regarding another exciting project of ours... We've run the first comm call of our series on the [statistical software review project](/stat-software-review/)! [rOpenSci Statistical Software Testing and Peer Review](/commcalls/feb2021-statsreview/) took place on March 2nd. The video recording (including closed captions) and resources as well as a summary have been [posted](/blog/2021/03/09/commcall-stats/).

Find out about more [events](/events).

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", toString(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", toString(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- toString(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the latest newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue)
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s.", name, url, description, user))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = toString(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    twitterImg = yaml$twitterImg %||% "",
    twitterAlt = yaml$twitterAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug
    )

  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, seq(nrow(posts)))
format_post <- function(post) {
  url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(post$date),
    stringr::str_pad(lubridate::month(post$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(post$date), 2, "0", side = "left"),
    post$slug
    )
  string <- sprintf("* [%s](%s) by %s", post$title, url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$twitterImg != "") {
    img_file <- fs::path_file(post$twitterImg)
    download.file(sprintf("https://ropensci.org/%s", post$twitterImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$twitterAlt)
    )
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, "software_peer_review")]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, "other")]
if (length(others) > 0) {
  cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, "tech_note")]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Citations

```{r cit}
citations <- jsonlite::read_json("https://ropensci-org.github.io/ropensci_citations/citations_all_parts_clean.json")
```

Below are the citations recently added to our database of `r length(citations)` articles, that you can explore on our [citations page](/citations).
We found use of...

```{r citations, results = "asis"}
since <- lubridate::as_date(last_newsletter) - 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci-org",
  repo = "ropensci_citations",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  )
)
old_commit <- gh::gh("/repos/{owner}/{repo}/commits/{ref}",
  owner = "ropensci-org",
  repo = "ropensci_citations",
  ref = commits[[length(commits)]]$parents[[1]]$sha
  )
old <- "https://raw.githubusercontent.com/ropensci-org/ropensci_citations/%s/citations_all_parts_clean.json" %>%
    sprintf(old_commit$sha) %>%
    jsonlite::read_json() 

new <- citations[length(old):length(citations)]

format_package <- function(package, packages = packages) {
  if (package %in% packages) {
    package <- sprintf("[**%s**](https://docs.ropensci.org/%s)", package, package)
  } else {
    package <- sprintf("**%s**", package)
  }
  package
}

format_one <- function(citation, packages) {
  packages <- toString(purrr::map_chr(citation$name, format_package, packages = packages))
  sprintf("* %s in %s\n\n", packages, citation$citation)
}

cat(
  paste0(
    sort(purrr::map_chr(new, format_one, packages = registry$package)),
    collapse = ""
  )
)

```

Thank you for citing our tools!

## Use cases

```{r usecases}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
usecases <- dplyr::filter(usecases, date >= anytime::anydate("2021-01-01"))#as.Date(last_newsletter))
usecases <- split(usecases, seq(nrow(usecases)))
```

`r snakecase::to_sentence_case(english(length(usecases)))` use cases of our packages and resources have been reported since the 1st of January.

```{r usecases2, results='asis'}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", usecase$title, usecase$url, usecase$reporter)
  string
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)!

## Call for maintainers

There's no open call for new maintainers at this point but you can refer to our [contributing guide](https://contributing.ropensci.org/) for finding ways to get involved!
As the maintainer of an rOpenSci package, feel free to contact us on Slack or email `info@ropensci.org` to get your call for maintainer featured in the next newsletter.

<!--IF CALL
* [our guidance on _Changing package maintainers_](https://devguide.ropensci.org/changing-maintainers.html)
* [our _Package Curation Policy_](https://devguide.ropensci.org/curationpolicy.html)

IF NO CALL
There's no open call for new maintainers at this point but you can refer to our [contributing guide](https://contributing.ropensci.org/) for finding ways to get involved!
As the maintainer of an rOpenSci package, feel free to contact us on Slack or email `info@ropensci.org` to get your call for maintainer featured in the next newsletter. -->

## Package development corner

Some useful tips for R package developers. :eyes:

Do you have some questions around **licensing of your R package**?
The [license chapter](https://r-pkgs.org/license.html) of the R packages book by Hadley Wickham and Jenny Bryan was [recently](https://twitter.com/hadleywickham/status/1307338435976364032) updated.
A less recent but interesting read is [Colin Fay's Licensing R e-book](https://thinkr-open.github.io/licensing-r/index.html).

**Encoding** can also be a tricky topic for package developers.
Useful resources include [Kevin Ushey's blog post "String encoding and R"](http://kevinushey.github.io/blog/2018/02/21/string-encoding-and-r/), [Irene Steves' blog post "Encoding in R"](https://irene.rbind.io/post/encoding-in-r/), searching for encoding in ["Writing R Extensions"](https://cran.r-project.org/doc/manuals/r-release/R-exts.html#Encoding) (e.g. "_If the DESCRIPTION file is not entirely in ASCII it should contain an ‘Encoding’ field specifying an encoding. This is used as the encoding of the DESCRIPTION file itself and of the R and NAMESPACE files, and as the default encoding of .Rd files._"), some [threads of R-package-devel](https://www.mail-archive.com/r-package-devel@r-project.org/msg06536.html).

Afraid to forget something when you **submit your package to CRAN**?
A check list could help!
You can use [`usethis::use_release_issue()`](https://usethis.r-lib.org/reference/use_release_issue.html), [number one "secret" [of the tidyverse team] for avoiding multiple rounds of review with CRAN](https://twitter.com/hadleywickham/status/1359852335330164737).
Now, if you want to add your own specific bullet e.g. ["Update my brag document"](https://jvns.ca/blog/brag-documents/) or ["Consider writing a tech note for rOpenSci blog"](https://blogguide.ropensci.org/index.html), you can add an [unexported `release_bullets()` function](https://github.com/r-lib/usethis/issues/1183) returning a character vector to your package.

```r
release_bullets <- function() {
  c(
    "Update my brag document.",
    "Contact rOpenSci blog editors to schedule a tech note."
  )
}
```

For inspiration you can search the [CRAN mirror maintained by R-hub](https://github.com/search?q=release_bullets+user%3Acran&type=Code).
For more resources around CRAN, see [Isabella Velásquez's post "You CRAN Do It, What CRAN First-Timers Should Know"](https://ivelasq.rbind.io/blog/you-cran-do-it/).

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Twitter account](https://twitter.com/ropensci).
