---
title: Test Katex
author:
  - Maëlle Salmon
date: '2023-01-26'
slug: test-katex
categories: []
tags:
  - Software Peer Review
  - packages
  - R
  - community
package_version: 0.1.0
description: A very short summary of your post (~ 100 characters)
twitterImg: blog/2019/06/04/post-template/name-of-image.png
twitterAlt: Alternative description of the image
tweet: A post about blabla by @username!
---

This is the R Markdown (.Rmd) template for a blog post or tech note. 
To generate your post with just Markdown (.md), use that template instead.

Throughout this template, including the YAML, 
you should change "post-template" to the slug of your post, 
and "2019-06-04" to your publication date.

Save this file under /content/blog/YYYY-MM-DD-slug/index.Rmd in the local copy of your roweb3 fork.

The chunk below ensures your plots will be inserted in the same folder as your Rmd so you can submit the bundle (.Rmd, .md, images) to roweb3.

```{r setup, include=FALSE}
# Options to have images saved in the post folder
# And to disable symbols before output
knitr::opts_chunk$set(fig.path = "", comment = "")

# knitr hook to make images output use Hugo options
knitr::knit_hooks$set(
  plot = function(x, options) {
    hugoopts <- options$hugoopts
    paste0(
      "{{<figure src=",
      '"', x, '" ',
      if (!is.null(hugoopts)) {
        glue::glue_collapse(
          glue::glue('{names(hugoopts)}="{hugoopts}"'),
          sep = " "
        )
      },
      ">}}\n"
    )
  }
)
```



`r katex::katex_html("y_1,\\ldots,y_T", displayMode = FALSE)`

```{r chunkname, hugoopts=list(alt="alternative text please make it informative", caption="this is what this image shows, write it here or in the paragraph after the image as you prefer", width=300)}
plot(1:10)
```

