---
slug: "news-september-2024"
title: rOpenSci News Digest, September 2024
author:
  - The rOpenSci Team
date: '2024-09-27'
tags:
  - newsletter
description: R-Universe list pages, training resources, coworking, new package and package news
params:
  last_newsletter: "2024-08-30"
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### Community call recording: Navigating the R ecosystem using R-universe

[Video and resources](/commcalls/nov2024-r-universe/).

Learn more about R-Universe and how you can use it to improve your R package development workflow.

In this community call, [Jeroen Ooms](/author/jeroen-ooms/) provided details on what R-Universe is and an update on what you can do with it today. 
He also discussed the future of R-Universe and how it can be used to navigate the R ecosystem.

### R-Universe list pages

R-Universe now features a [list of all datasets in all packages](https://r-universe.dev/datasets) and a [list of all vignettes  from CRAN, Bioconductor, and others](https://r-universe.dev/articles).

Another important global table is the [package scoreboard](https://r-universe.dev/scores) that lets you explore the scores used by R-Universe search engine for ranking, which is based on:

- Stars: Number of stars on GitHub;
- Downloads per month from CRAN or Bioconductor mirrors;
- Scripts: (NEW) number of files on GitHub that mention 'library(pkgname)';
- Dependents: recursive reverse dependencies;
- Unique contributors;
- Yearly commits.

### Resources from rOpenSci training sessions

We've added a searchable table of [past training sessions](/training/) to our resources hosted on the rOpenSci website.
Explore the full list of our [resources](/resources).

### Noam Ross, rOpenSci, and communities empowering open science

DataLatam Podcast [interviewed](https://www.datalatam.com/109/) rOpenSci director Noam Ross. 
Learn about his work at the intersection of health and the environment and his journey with rOpenSci.

### Materials: Screen Reader Accessible Tools and Resources for Learning and Working with R

We have now published the recording of the webinar that walks through learning and using R with screen readers — and it's bilingual (English & Turkish)! with subtitles in English.

📹 [Webinar video](https://vimeo.com/1008631708) featuring [Liz Hare, PhD](/author/liz-hare/)
and [Alican Cagri Gokcek](/author/alican-cagri-gokcek/)

Also, don't miss the [detailed technical note](/blog/2024/09/05/screen-readers-tools/) by Liz Hare with all the necessary resources to get started.

### The rOpenSci community at upcoming events

rOpenSci Community Manager [Yanina Bellini Saibene](/author/yanina-bellini-saibene/) will deliver a keynote talk at [BioNT Community Event & CarpentryConnect-Heidelberg 2024](/events/carpentryconnect-2024/), on November 14th in Heidelberg, Germany.

[Will Landau](/author/will-landau/) will deliver a keynote talk at [LatinR (Latin American Conference About the Use of R in R&D)](/events/latinr-2024/). The conference is happening online from November 18-22. 


### Coworking

Read [all about coworking](/blog/2023/06/21/coworking/)!

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.

- [Tuesday, October 1st, 14h00 Europe Central (12:00 UTC)](/events/coworking-2024-10), "Using R to tell (data science) stories" with cohost [Apoorv Anand](/author/apoorv-anand) and [Steffi LaZerte](/author/steffi-lazerte).
    - Explore how others use R to tell data stories
    - Start work on a project using R to craft a narrative through data
    - Chat with our cohost to discuss and explore the use of R in data stories
- [Tuesday, November 5th, 09:00 Americas Pacific (17:00 UTC)](/events/coworking-2024-11), "Getting to know Openscapes" with cohost [Stefanie Butland](/author/stefanie-butland) and [Steffi LaZerte](/author/steffi-lazerte).
    - Look up [Openscapes](https://openscapes.org/)
    - Chat with our cohost to learn more about the Openscapes organization, and how they use R to automate creating open education infrastructure

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE,eval=TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (package$package == "karel") {
      rev <- c("Veronica Jimenez-Jacinto", "Joel Nitta")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md",  "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$socialImg != "") {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt)
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).

* [medrxivr](https://docs.ropensci.org/medrxivr/), access and search MedRxiv and BioRxiv preprint data. [Issue for volunteering](https://github.com/ropensci/medrxivr/issues/41).

* [USAboundaries](https://docs.ropensci.org/USAboundaries/) (and USAboundariesdata), historical and contemporary boundaries of the United States of America . [Issue for volunteering](https://github.com/ropensci/USAboundaries/issues/50).

* [historydata](https://docs.ropensci.org/historydata/), datasets for historians. [Issue for volunteering](https://github.com/ropensci/historydata/issues/23).

### Calls for contributions

* [Help make qualtRics better! Code contributions wanted](/blog/2024/03/01/qualtrics-call-help/)

* [Help make assertr better! Come close issues](/blog/2024/02/27/assertr-call-help/)

* [Help users and developers of targets by answering questions!](/blog/2024/02/29/targets-call-help/)

* [Help make waywiser better! User requests wanted](/blog/2024/02/19/waywiser-call-help/)

Also refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.

## Package development corner

Some useful tips for R package developers. :eyes:

### Reminder: R Consortium Infrastructure Steering Committee (ISC) Grant Program Accepting Proposals until October 1st!

The R Consortium [Call for Proposals](https://www.r-consortium.org/all-projects/call-for-proposals) might be a relevant funding opportunity for your package!
If you can't prepare your proposal in time, the next call should start [March 1st 2025](https://www.r-consortium.org/all-projects/call-for-proposals).

### How to demonstrate an error in an R Markdown vignette

An useful piece of information shared by Hugo Gruson in our [Slack workspace](https://contributing.ropensci.org/resources.html#channels).
To demonstrate an error in a vignette, try setting `purl = FALSE` in chunks with `error = TRUE`.
[More info in a knitr issue](https://github.com/yihui/knitr/issues/2338).

### Better titles for roxygen2 "family" cross-references

Do you use the "family" roxygen2 tag for grouping and cross-linking manual pages?
Did you know you can override the default title from, for instance, "Other blabla" to "blabla functions and related data"?
If not, refer to [roxygen2 documentation](https://roxygen2.r-lib.org/articles/index-crossref.html#family).

### Create and use a custom roxygen2 tag

Read more about creating a custom tag that adds a new section to a manual page in this [blog post by Maëlle Salmon](https://masalmon.eu/2024/09/03/roxygen2-custom-tag/).

### Materials from posit::conf(2024) workshops

Materials from posit::conf(2024) workshops have been [posted](https://posit-conf-2024.github.io/workshops/).
Especially relevant to package developers: [From R User to R Programmer](https://posit-conf-2024.github.io/programming-r/) by Emma Rand and Ian Lyttle; and [Package Development: The Rest of the Owl](https://github.com/posit-conf-2024/pkg-dev) by Jenny Bryan whose introduction slidedeck features GitHub search tips.

### Some tips when searching GitHub issues

Speaking of GitHub search...

Michael Chirico shared the following tip on Mastodon: ["[you can use] the 'in:title' operator in GitHub search to restrict a search to issues/PRs matching a query in the title"](https://fosstodon.org/@michaelchirico/113083595380080927).

Jenny Bryan answered with another tip: ["involves:USERNAME" for "I was once part of a conversation about this"](https://mastodon.social/@jennybryan@fosstodon.org/113085363392572924).

### A usethis function to create a test helper file

The new usethis function [` usethis::use_test_helper("mocks")`](https://usethis.r-lib.org/reference/use_test_helper.html) will create a [helper file](https://testthat.r-lib.org/articles/special-files.html) under `tests/testthat/helper-mocks.R`.

### A devtools function for building the README and vignettes

Not a new function, but in case you missed it...
The [`devtools::build_rmd()` and `devtools::build_readme()` functions](https://devtools.r-lib.org/reference/build_rmd.html) render R Markdown files using a temporary install of your current development package.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
