---
title: rOpenSci News Digest, March 2023
author:
  - The rOpenSci Team
date: '2023-03-17'
slug: ropensci-news-digest-march-2023
categories: []
tags:
  - newsletter
description: keywords from the content
params:
  last_newsletter: '2023-02-17'
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  glue::glue_collapse(..., sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### Meeting the stars of the R-universe: Sébastien Rochette

Knowing our community's stories helps us to learn about the people behind our software, brings us closer and offers us new opportunities. To share some of these community stories, we created the rOpenSci interview series [_"Meeting the stars of the R-Universe"_](/tags/r-universe-stars/).

The latest interview with [Sébastien Rochette introduces ThinkR’s Approach to Contributing to a Growing and Friendly R Community](/blog/2023/02/28/r-universe-stars-2-en/).
The post is available in [Spanish](/blog/2023/02/28/r-universe-stars-2-es/) and [French](/blog/2023/02/28/r-universe-stars-2-fr/) too!
Don't miss the trilingual post and the video.


### Discovering and learning everything there is to know about R packages using R-universe

Jeroen Ooms explains [how to use R-universe to discover and assess new packages](/blog/2023/02/27/runiverse-discovering/).
He wrote that we can distinguish three levels of navigation in the R-universe when you go shopping for R packages:

1. Search the global ecosystem: find packages, by topic, keyword, ranking, etc.
1. Browse by maintainer/organization: explore all work from a given group or developer.
1. The individual package: get detailed information on everything there is to know about a project and instructions for how to start using it.

That post was also discussed on the [R Weekly highlights podcast](https://rweekly.fireside.fm/114?t=576) hosted by Eric Nantz and Mike Thomas!

### Coworking

Join us for social coworking & office hours monthly on first Tuesdays! 
Hosted by Steffi LaZerte and various community hosts. 
Everyone welcome. 
No RSVP needed. 
Consult our [Events](/events) page to find your local time and how to join.
    
* [Tuesday, Apr 4th, 14:00 European Central / 12:00 UTC](/events/coworking-2023-04/) "Working with taxonomic lists in R" *Hosted by community host [Miguel Alvarez](/author/miguel-alvarez/) and [Steffi LaZerte](/author/steffi-lazerte/)*
    - Explore R packages for working with lists (such as [taxlist](https://docs.ropensci.org/taxlist/)),
    - Organize some taxonomy in your data using taxlist,
    - Talk to [Miguel](/author/miguel-alvarez/) and discuss tips for working with taxonomic lists.
    
And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

* Tuesday, May 2nd, 9:00 Americas Pacific / 16:00 UTC Tentative theme: "Spring Cleaning for R packages and scripts" *Hosted by community host TBD and [Steffi LaZerte](/author/steffi-lazerte/)*
    - Explore how other organizations keep their scripts/packages nice and clean
    - Take a look at your R packages and scripts and give them a good spring cleaning\*
    - Talk to our community host and other attendees and discuss tips for keeping on top of it all.
    
\* in the northern hemisphere at least, otherwise, give them a good *fall* cleaning!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% c("babeldown", old$package),
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE}
packages <- split(new, seq(nrow(new)))
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(x$given, x$family))
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login
  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state, 
    labels = label
  )
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- paste0("https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A", label)
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))
  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    twitterImg = yaml$twitterImg %||% "",
    twitterAlt = yaml$twitterAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug
    )

  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md", "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, seq(nrow(posts)))
format_post <- function(post) {
  url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(post$date),
    stringr::str_pad(lubridate::month(post$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(post$date), 2, "0", side = "left"),
    post$slug
    )
  string <- sprintf("* [%s](%s) by %s", post$title, url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("\\?$", "", sub("\\!$", "", sub("\\.$", "", post$description), ".")), ".")
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$twitterImg != "") {
    img_file <- fs::path_file(post$twitterImg)
    download.file(sprintf("https://ropensci.org/%s", post$twitterImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    string <- paste0(
      string,
      sprintf('{{< figure src="%s" alt="%s" width="400" >}}\n\n', img_file, post$twitterAlt)
    )
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, "software_peer_review")]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, "other")]
if (length(others) > 0) {
  if (length(others) != length(posts)) cat("### Other topics\n\n")
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, "tech_note")]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

<!-- ## Use cases

```{r usecases, eval=FALSE}
# rerun get_use_cases.R at the same time
usecases <- jsonlite::read_json("../../../data/usecases/usecases.json")
get_one_case <- function(usecase) {
  tibble::tibble(
    title = usecase$title,
    reporter = usecase$reporter,
    url = usecase$url,
    image = usecase$image,
    date = anytime::anydate(usecase$date)
  )
}
usecases <- purrr::map_df(usecases, get_one_case)
usecases <- dplyr::filter(usecases, date >= as.Date(last_newsletter))
usecases <- split(usecases, seq(nrow(usecases)))
```

`snakecase::to_sentence_case(english(length(usecases)))` use case` if (length(usecases) > 1) "s"` of our packages and resources ha` if (length(usecases) > 1) "ve" else "s"` been reported since we sent the last newsletter.

```{r usecases2, results='asis', eval=FALSE}
format_case <- function(usecase) {
  string <- sprintf("* [%s](%s). Reported by %s.", sub("\\.$", "", usecase$title), usecase$url, usecase$reporter)
}
cat(
  paste0(
    purrr::map_chr(usecases, format_case),
    collapse = "\n\n"
  )
)
```

Explore [other use cases](/usecases) and [report your own](https://discuss.ropensci.org/c/usecases/10)! -->

## Call for (co)maintainers

### Call for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/) (or listening to its discussion on the [R Weekly highlights podcast](https://rweekly.fireside.fm/111) hosted by Eric Nantz and Mike Thomas)!

- **[rvertnet](https://cran.r-project.org/web/packages/rvertnet/index.html)**, Retrieve, map and summarize data from the VertNet.org archives (<http://vertnet.org/>). Functions allow searching by many parameters, including taxonomic names, places, and dates. In addition, there is an interface for conducting spatially delimited searches, and another for requesting large datasets via email. [Issue for volunteering](https://github.com/ropensci-archive/rvertnet/issues/71).

- **[natserv](https://cran.r-project.org/web/packages/natserv/index.html)**. Interface to NatureServe (<https://www.natureserve.org/>). Includes methods to get data, image metadata, search taxonomic names, and make maps. [Issue for volunteering](https://github.com/ropensci-archive/natserv/issues/29).

- **[sofa](https://cran.r-project.org/web/packages/sofa/index.html)**. Provides an interface to the NoSQL database CouchDB (<http://couchdb.apache.org>). Methods are provided for managing databases within CouchDB, including creating/deleting/updating/transferring, and managing documents within databases. One can connect with a local CouchDB instance, or a remote 'CouchDB' databases such as Cloudant. Documents can be inserted directly from vectors, lists, data.frames, and JSON. Targeted at CouchDB v2 or greater. [Issue for volunteering](https://github.com/ropensci-archive/sofa/issues/81).

- **[geojsonlint](https://cran.r-project.org/web/packages/geojsonlint/index.html)**, Tools for linting GeoJSON. Includes tools for interacting with the online tool <http://geojsonlint.com>, the Javascript library geojsonhint (<https://www.npmjs.com/package/geojsonhint>), and validating against a GeoJSON schema via the Javascript library (<https://www.npmjs.com/package/is-my-json-valid>). Some tools work locally while others require an internet connection. [Issue for volunteering](https://github.com/ropensci-archive/geojsonlint/issues/22).

- **[citesdb](https://docs.ropensci.org/citesdb/)**, a high-performance database of shipment-level CITES trade data. Provides convenient access to over 40 years and 20 million records of
endangered wildlife trade data from the Convention on International Trade
in Endangered Species of Wild Fauna and Flora, stored on a local on-disk,
out-of memory 'DuckDB' database for bulk analysis. [Issue for volunteering](https://github.com/ropensci/citesdb/issues/21).


### Call for comaintainers

- **[rtweet](https://docs.ropensci.org/rtweet)**, that interfaces Twitter API, is looking for a [co-maintainer](https://github.com/ropensci/rtweet/issues/763).

- Refer to our [recent blog post](/blog/2022/10/17/maintain-or-co-maintain-an-ropensci-package/#packages-looking-for-co-maintainers) to identify other packages where help is especially wished for!

## Package development corner

Some useful tips for R package developers. :eyes:

### R Consortium's call for proposals!

The R Consortium's Internal Steering Committee has a [call for proposals](https://www.r-consortium.org/all-projects/call-for-proposals) open until April 1st.

- The funds can be used for different sizes of projects. The project must have a software development component.
- Proofs of concept are not funded. No scientific publications or equipment.
- The idea is that the funds can cover people's time to develop software.

This might be relevant for your R package work so make sure to read the call, and good luck if you send a proposal! :rocket:

### To cache, or not to cache testthat results?

Have you ever wished you could cache testthat results?
You'll find arguments both in favor of and against that idea in this [testthat issue](https://github.com/r-lib/testthat/issues/1703) -- testthat maintainer Hadley Wickham being against the idea.

You might be interested in Kirill Müller's experimental package [lazytest](https://krlmlr.github.io/lazytest/) that helps you rerun only the tests that have failed during the last run.

### Check if an R package name is available 

The function [`pak::pkg_name_check()`](https://pak.r-lib.org/reference/pkg_name_check.html) by Gábor Csárdi can be viewed as a replacement for the available package. It has a very nice output.

### What if your httptest mock files are suddenly ignored?

Imagine you've set up [HTTP testing](https://books.ropensci.org/http-testing/index.html) in your package with [httptest](https://books.ropensci.org/http-testing/httptest.html) and all goes well until one day, where the httptest mock files are _ignored_.
Don't panick! 
Check whether the calls that are mocked are still made with httr.
Maybe one of your package's dependencies upgraded their [stack](https://books.ropensci.org/http-testing/http-in-r-101.html#http-requests-in-r-what-package)?
If the calls are made with httr2, the tests need to be updated to [httptest2](https://books.ropensci.org/http-testing/httptest2.html) which thankfully isn't too hard.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).
