---
title: A better R programming experience thanks to tree-sitter
author: 
- Maëlle Salmon
editor: 
date: '2026-03-26'
slug: tree-sitter-overview
description: Modern tooling for parsing, searching, formatting, editing R code, just like for other programming languages.
output: hugodown::md_document
tags:
  - package development
  - tech notes
params:
  doi: "10.59350/57mzc-7e212"
output: hugodown::md_document
---

A little bit less than two years ago, building on work by Jim Hester and Kevin Ushey, Davis Vaughan completed a very impactful JSON file for the R community:
a grammar for R for the tree-sitter parsing generator.
He even got a round of applause for it during a talk at the useR! 2024 conference!
So, did he get cheered for... JSON? :sweat_smile:
No, the audience got excited about the _improved developer experience for R_ that this file unlocked. 
In this post, we'll explain what tree-sitter is, and what tools built upon it can bring to your workflow writing R software.

## Code parsing: what is tree-sitter

[Tree-sitter](https://en.wikipedia.org/wiki/Tree-sitter_(parser_generator)) is a code [parsing generator](https://tree-sitter.github.io/tree-sitter/) with bindings existing in several languages including Rust. 
It is written in C, the CLI is written in Rust, parsers are generated from JSON files [using Node](https://github.com/tree-sitter/tree-sitter/blob/master/docs/src/6-contributing.md).

Let's rewind a little bit.
What does it mean to parse code?

Basically, given a string of code like

```r
a <- mean(x, na.rm = TRUE)
```

How do you know that `mean` is a function name, `na.rm` an argument name, `TRUE` a logical?
You have to _parse_ that code into what's called a parse tree.
You do that in your head when reading R code. :smile_cat:

R itself can obviously parse R code, you can use `parse()` and `getParseData()` to reproduce that behavior.

```{r}
parse(
  text = "a <- mean(x, na.rm = TRUE)",
  keep.source = TRUE
) |>
  getParseData()
```

Or you could transform that same data into XML using [{xmlparsedata}](https://r-lib.github.io/xmlparsedata/):

```{r}
parse(
  text = "a <- mean(x, na.rm = TRUE)",
  keep.source = TRUE
) |>
  xmlparsedata::xml_parse_data(pretty = TRUE) |>
  xml2::read_xml() |>
  as.character() |>
  cat()
```

In both cases, you recognize words such as `LEFT_ASSIGN` or `SYMBOL_FUNCTION_CALL`.
Parsing is useful to execute code if you're R,
and otherwise to analyse code (does it call such and such function? and this without your own brittle regular expressions), to navigate it (going from a function call to the definition of that function), to modify it (replacing all occurrences of a function with another one).

Now, tree-sitter allows such code parsing but **faster** especially thanks to its support of incremental parsing -- updating the syntax tree as you are typing in your editor for instance!
Tree-sitter is agnostic in that it can parse any code as long as it gets the Rosetta stone for it. 
It's been used for many languages which means many tools have been built around it.

To make tree-sitter "learn" a new language you need to give it JSON containing the definition of the syntax of that language,
what's called a _grammar_.
This is where the aforementioned JSON file by Davis Vaughan and collaborators comes into play!
Treesitter.r, that provides the R grammar, is key to all tools presented in this post. 
The grammar is [data](https://github.com/r-lib/tree-sitter-r/blob/main/src/grammar.json) that can be used by other tools, not only R packages, even if the grammar is among other things shared as an R package called treesitter.r.


## Browsing code interactively: Positron IDE, GitHub

## Searching code: ast-grep

## Formatting and linting: Air, Jarl, flir

## Testing code: muttest

https://github.com/jakubsob/muttest

## Diffing code: difftastic

## Conclusion: more to come?

Modern

Ecosystem in flux
