---
slug: "news-november-2025"
title: rOpenSci News Digest, November 2025
author:
  - The rOpenSci Team
date: '2025-11-28'
tags:
  - newsletter
description: LatinR, dev guide in Portuguese, posit::conf(2025) recordings, coworking, new package, package news
params:
  last_newsletter: '2025-10-24'
  doi: "10.59350/7pygv-c1m20"
---

```{r setup, include=FALSE}
library("magrittr")
library("rlang")
last_newsletter <- anytime::anytime(params$last_newsletter)
knitr::opts_chunk$set(echo = FALSE)
url <- sprintf(
    "/blog/%s/%s/%s/%s",
    lubridate::year(rmarkdown::yaml_front_matter(knitr::current_input())$date),
    stringr::str_pad(lubridate::month(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(rmarkdown::yaml_front_matter(knitr::current_input())$date), 2, "0", side = "left"),
    rmarkdown::yaml_front_matter(knitr::current_input())$slug
    )
english <- function(x) {
  as.character(english::english(x))
}

nice_string <- function(...) {
  if (length(...) == 2) {
    return(paste(c(...), collapse = " and "))
  }
  glue::glue_collapse(trimws(...), sep = ", ", last = ", and ")
}
```
<!-- Before sending DELETE THE INDEX_CACHE and re-knit! -->

Dear rOpenSci friends, it's time for our monthly news roundup!
<!-- blabla -->
You can read this post [on our blog](`r url`).
Now let's dive into the activity at and around rOpenSci!

## rOpenSci HQ

### rOpenSci at LatinR

We’re excited to continue supporting LatinR as a community partner in 2025. [Registration is now open for the free LatinR Conference](https://www.eventbrite.com.ar/e/1939670499679), bringing together researchers, developers, and open science advocates from across the region.

#### Keynotes

* Heather Turner — Lowering Barriers to Contributing to R
* Stephanie Zimmer — Transforming a team to open-source first
* TRACE-LAC Team — Lo invisible del código abierto: Lessons from the TRACE-LAC / Epiverso project for connecting software development with public health

#### Talks in Spanish

- **Herramientas para usar LLMs en R** by [Luis D. Verde Arregoitia](/author/luis-d.-verde-arregoitia/)
- **R-multiverse** by [Will Landau](/author/will-landau/), [Maëlle Salmon](/author/maëlle-salmon/) and [Yanina Bellini Saibene](/author/yanina-bellini-saibene/)
- **Mejor código, sin esfuerzos, sin siquiera IA** by [Maëlle Salmon](/author/maëlle-salmon/), [Hugo Gruson](/author/hugo-gruson/) and Etienne Bacher 
- **Estrategias de divulgación para proyectos de software e infraestructuras abiertas** by [Alejandra Bellini](/author/alejandra-bellini/) and [Yanina Bellini Saibene](/author/yanina-bellini-saibene/)
- **Comunidades de líderes de código abierto** by [Yanina Bellini Saibene](/author/yanina-bellini-saibene/) and [Noam Ross](/author/noam-ross/)
- **Datos públicos y Software Libre** by [Pablo Tiscornia](/author/pablo-tiscornia/)
- **rcdo para analizar datos climáticos con R** by [Elio Campitelli](/author/elio-campitelli/)
- **ICC para evaluar confiabilidad entre evaluadores** by [Francisco Cardozo](/author/francisco-cardozo/)
- **Integrando listas taxonómicas en Quarto y R Markdown: Un caso para `taxnames`** by [Miguel Alvarez](/author/miguel-alvarez/)
- **Metasurvey** by [Mauro Loprete](/author/mauro-loprete/).

#### Talks in English

- **R-universe Q&A** [Jeroen Ooms](/author/jeroen-ooms/) and [Maëlle Salmon](/author/maëlle-salmon/)

#### Tutorials in Spanish

- **¡Miércoles, Git! Manejo de errores en Git y no morir en el intento** — [Maëlle Salmon](/author/maëlle-salmon/) & [Yanina Bellini Saibene](/author/yanina-bellini-saibene/). Tuesday, December 2. 10:00-12:00 UTC-3.
- **Introducción a Tidymodels** — [Francisco Cardozo](/author/francisco-cardozo/) & Edgar Ruiz. Monday, December 1. 18:00-20:00 UTC-3.
- **Automatización de workflows en R y Python con targets y snakemake** — [Diana Garcia](/author/diana-garcia/). Tuesday, December 2. 14:00-16:00 UTC-3.
- **¿Qué historia vas a contar hoy? Herramientas para una comunicación eficaz** — [Alejandra Bellini](/author/alejandra-bellini/). Tuesday, December 2. 17:00-19:15 UTC-3.

#### Tutorials in English

- **Coding with AI in RStudio** — Juan Cruz Rodríguez & [Luis D. Verde Arregoitia](/author/luis-d.-verde-arregoitia/). Tuesday, December 2. 10:00-12:00 UTC-3.

Full tutorial schedule: https://latinr.org/en/cronograma/tutoriales/workshops.html

### rOpenSci Dev Guide Now Available in Portuguese!

rOpenSci has published the Portuguese translation of the “Packages: Development, Maintenance, and Peer Review” guide, expanding access to best practices in R package development across the Portuguese-speaking world.

The two-year, community-led effort brought together contributors from Angola, Brazil, and Portugal, following rOpenSci’s open workflow with automated translation, multi-reviewer editing, and collaborative decisions on terminology and inclusivity.

This new version strengthens rOpenSci’s multilingual infrastructure and helps lower language barriers for contributors.

Read the full [blog post to learn more about the project](/blog/2025/11/25/translation-devguide-pt/).

### Big Team Collaboration on Data Management and Analysis with rOpenSci

[Liz Hare](/author/liz-hare) and [Yanina Bellini Saibene](/author/yanina-bellini-saibene/) gave a talk on Big Team Collaboration on Software Peer Review with rOpenSci at the Big Team Science Conference 2025.
The slides and recording have been [posted](/events/BTSConf-2025)!

### rOpenSci at posit::conf(2025): recordings

Recordings from posit::conf(2025) are now [publicly available](https://posit.co/blog/talks-and-workshops-from-posit-conf-2025/) on YouTube!
Many excellent talks to enjoy and learn from, including those from rOpensci community members.

* [Mauro Lepore](/author/mauro-lepore) and [Eric Scott](/author/eric-scott) both presented lightning talks. Mauro talked about ["Approaching Positron from RStudio"](https://www.youtube.com/watch?v=X6PZ_RiZp44), Eric about ["Enabling geospatial workflow management with targets: an R package origin story"](https://www.youtube.com/watch?v=RmTtugX1CoE).

* [Will Landau](/author/will-landau/) gave his talk ["R-multiverse: a next-generation R package repository system built on R-universe"](https://www.youtube.com/watch?v=UseltnRVoFo).

* [Nic Crane](/author/nic-crane) presented on ["Hacking Productivity with LLMs: What Works (and What Doesn’t)"](https://www.youtube.com/watch?v=_nhDYt0f5yc). 

* [Noam Ross](/author/noam-ross/) and [Beatriz Milz](/author/beatriz-milz/) both presented in the session "It Takes a Village: Building and Sustaining Communities". Beatriz talked about ["Translating R for Data Science into Portuguese: A Community-Led Initiative"](https://www.youtube.com/watch?v=yBC8rNvh3Ns), Noam about ["rOpenSci Champions: Building Communities of Open-Source Leaders"](https://www.youtube.com/watch?v=-atDblWsOBU).

- [Luis D. Verde Arregoitia](https://ropensci.org/author/luis-d.-verde-arregoitia/) presented his talk ["Bold indicates negative?"](https://www.youtube.com/watch?v=NJz0779Seto).

### Blog contributor roles in rOpenSci and beyond

We recently added support for contributor roles on our [blog](/blog/2025/10/14/blog-roles/): author, editor, translator, interviewee.
Rogue Scholar, that generates the DOI of our posts (and many others'), also added this feature.
Read more in their [announcement](https://blog.front-matter.de/posts/supporting-blog-contributions-beyond-authorship/).
Thanks to Martin Fenner for supporting our efforts in recognizing all contributions to blog posts!

### Coworking 

Read [all about coworking](/blog/2023/06/21/coworking/)!

-   Tuesday December 2nd, 14h00 Europe Central (13:00 UTC), ["Getting to know The Carpentries"](/events/coworking-2025-12/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Angelique Trusler](/author/angelique-trusler/).  
    -   Visit [The Carpentries](https://carpentries.org/);  
    -   Meet community host, Angelique Trusler, and learn more about The Carpentries and how you might get involved. 
-   Tuesday January 13th, 9:00 Americas Pacific (17:00 UTC), ["Let it go!"](/events/coworking-2026-01/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Yanina Bellini Saibene](/author/yanina-bellini-saibene/).  
    -   Spend some time reviewing the forums, Slack workspaces, Newsletters, RSS feeds (etc. etc.) you’re subscribed to;
    -   Unsubscribe to all you no longer need (Let it go!);
    -   Meet co-host, Yanina Bellini Saibene, and discuss strategies for this New Year decluttering of your digital (or perhaps not-so-digital) life.
-   Tuesday February 2nd, 9:00 Australia Western (01:00 UTC), ["Share your Positron setup!"](/events/coworking-2026-02/) with [Steffi LaZerte](/author/steffi-lazerte/) and cohost [Noam Ross](/author/noam-ross/).  
    -   Setup Poistron and explore extensions and custom settings;
    -   Meet community host, Noam Ross, share how you have set up Positron for your workflow and learn from others.

And remember, you can always cowork independently on work related to R, work on packages that tend to be neglected, or work on what ever you need to get done!

## Software :package:

### New packages

```{r new-packages, cache = TRUE}
cran_unquote <- function(string) {
  gsub("\\'(.*?)\\'", "\\1", string)
}
tidy_package <- function(entry) {
  tibble::tibble(
    package = entry$name,
    description = cran_unquote(entry$description),
    details = cran_unquote(entry$details),
    on_cran = entry$on_cran,
    on_bioc = entry$on_bioc,
    onboarding = entry$onboarding,
    url = entry$url,
    maintainer = entry$maintainer # use desc for more info
    
  )
}

registry <- "https://raw.githubusercontent.com/ropensci/roregistry/gh-pages/registry.json" %>%
  jsonlite::read_json() %>%
  purrr::pluck("packages") %>%
  purrr::map_df(tidy_package)
  
since <- lubridate::as_date(last_newsletter) - 1
until <- lubridate::as_date(last_newsletter) + 1
commits <- gh::gh(
  "GET /repos/{owner}/{repo}/commits",
  owner = "ropensci",
  repo = "roregistry",
  since = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(since),
    stringr::str_pad(lubridate::month(since), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(since), 2, "0", side = "left")
  ),
  until = sprintf(
    "%s-%s-%sT00:00:00Z",
    lubridate::year(until),
    stringr::str_pad(lubridate::month(until), 2, "0", side = "left"),
    stringr::str_pad(lubridate::day(until), 2, "0", side = "left")
  )
)

empty <- TRUE
i <- length(commits)
while (empty == TRUE) {
  old <- "https://raw.githubusercontent.com/ropensci/roregistry/%s/packages.json" %>%
    sprintf(commits[[i]]$sha) %>%
    jsonlite::read_json() %>%
    purrr::map_df(function(x) tibble::tibble(package = x$package, url = x$url, branch = x$branch))
  i <- i - 1
  if (nrow(old) > 100) {
    empty <- FALSE
  }
}

old <- dplyr::filter(
  old,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

new <- dplyr::filter(
  registry,
  !package %in% old$package,
  !package %in% "gpg",
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)
```


The following `r if(nrow(new)>1) english(nrow(new))` package`r if(nrow(new)>1) "s"` recently became a part of our software suite:

```{r, results='asis', cache = TRUE,eval=TRUE}
new$onboarding[new$package == "hdcuremodels"] <- "https://github.com/ropensci/software-review/issues/692"
packages <- split(new, seq(nrow(new)))
packages <- rev(packages)
present_one <- function(package) {
  url_parts <- urltools::url_parse(package$url)
  desc_link <- gh::gh(
    "/repos/{owner}/{repo}/contents/{path}",
    owner = strsplit(url_parts$path, "\\/")[[1]][1],
    repo = strsplit(url_parts$path, "\\/")[[1]][2],
    path = "DESCRIPTION"
  ) %>%
    purrr::pluck("download_url")
  
  if (package$package == "eDNAjoint") {
    desc_link <- "https://raw.githubusercontent.com/ropensci/eDNAjoint/26fb2d22dcd4314cdf7d170812c6247ae51f25a0/DESCRIPTION"
  }
  
  withr::with_tempfile(
    "tf", {
      download.file(desc_link, tf) 
      desc <<- desc::desc(file = tf)
    }
  )
  # as in pkgdown
  authors <- unclass(desc$get_authors())
  aut <- purrr::keep(authors, function(x) {any( x$role %in% "aut") && all(x$role != "cre") })
  aut <- purrr::map_chr(aut, function(x) paste(x$given, x$family))
  rev <- purrr::keep(authors, function(x) {any( x$role %in% "rev") && all(x$role != "cre") })
  rev <- purrr::map_chr(rev, function(x) paste(c(x$given, x$family), collapse = " "))
  if (package$package == "forcis") {
    rev <- c(rev, "Scott Chamberlain")
  }
  if (package$package == "sasquatch") {
    rev <- c("Nic Crane")
  }
  maintainer <- purrr::keep(authors, function(x) {any( x$role %in% "cre") })
  maintainer <- paste(c(maintainer[[1]]$given, maintainer[[1]]$family), collapse = " ")
  author_string <- sprintf("developed by %s", maintainer)
  
  if (length(aut) > 0) {
    author_string <- paste0(author_string, sprintf(" together with %s", nice_string(aut)))
  } 
  
  string <- sprintf(
    "[%s](https://docs.ropensci.org/%s), %s: %s. ",
    package$package, 
    package$package, 
    author_string,
    stringr::str_remove(stringr::str_squish(package$details), "\\.$")
  )

  string <- gsub("doi\\:", "https://doi.org/", string)
  
  if (package$on_cran) {
    string <- paste0(
      string, 
      sprintf(
        " It is available on [CRAN]( https://CRAN.R-project.org/package=%s). ",
        package$package
      )
    )
  }
  if (package$on_bioc) {
    string <- paste0(
      string, sprintf(
        " It is available on [Bioconductor](https://bioconductor.org/packages/%s/). ",
        package$package
      )
    )
  }
  if (nzchar(package$onboarding)) {
    string <- paste0(string, sprintf("It has been [reviewed](%s)", package$onboarding))
    if (package$package == "fireexposuR") {
      rev <- c("Sherry Zhang ", "Ronny A. Hernández Mora")
    }
    if (length(rev) > 0) {
      string <- paste0(string, sprintf(" by %s.", nice_string(rev)))
    } else {
      string <- paste0(string, ".")
    }
  }
  string <- sub("twitter", "social media", string)
  paste("+", string)

}
text <- purrr::map_chr(
  packages,
  present_one
)
cat(paste0(text, collapse = "\n\n"))
```

Discover [more packages](/packages), read more about [Software Peer Review](/software-review).

### New versions

```{r news, cache=TRUE}
registry <- dplyr::filter(
  registry,
  !grepl("ropenscilabs\\/", url),
  !grepl("ropensci-archive\\/", url)
)

registry <- registry %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
  owner = strsplit(urltools::path(url), "/")[[1]][1],
  repo = strsplit(urltools::path(url), "/")[[1]][2]
) %>%
  dplyr::filter(
    !is.na(owner)
  )
packages <- split(registry, seq(nrow(registry)))
get_release <- function(repo) {
  info <- gh::gh(
    "GET /repos/{owner}/{repo}/releases",
    owner = repo$owner,
    repo = repo$repo,
    per_page = 3,
    description = repo$description
  )
  info <- info[!purrr::map_lgl(info, "draft")]
  if(length(info) == 0 || anytime::anytime(info[[1]]$published_at) < last_newsletter) {
    return(NULL)
  }
  
  tibble::tibble(
    package = repo$package,
    version = info[[1]]$tag_name,
    url = info[[1]]$html_url,
    description = repo$description
  )
}
releases <- purrr::map_df(
  packages,
  get_release
)
releases <- split(releases, seq(nrow(releases)))
format_release <- function(release) {
  sprintf(
    '[%s](https://docs.ropensci.org/%s "%s") ([`%s`](%s))',
    release$package,
    release$package,
    release$description,
    release$version,
    release$url
  )
}
all_releases <- purrr::map_chr(releases, format_release)
text <- nice_string(all_releases)
```

The following `r if (length(releases) > 1) english(length(releases))` package`r if (length(releases) > 1) "s"` `r if (length(releases) > 1) "have" else "has"` had an update since the last newsletter: `r text`.

## Software Peer Review

```{r software-review, results='asis'}
# from pkgdown https://github.com/r-lib/pkgdown/blob/1ca166905f1b019ed4af9642617ea09fa2b8fc17/R/utils.r#L176

get_description <- function(body) {
  lines <- strsplit(body, "\n")[[1]]
  name <- stringr::str_squish(sub("Package:", "", lines[grepl("^Package", lines)][1]))
  description <- stringr::str_squish(sub("Title:", "", lines[grepl("^Title", lines)][1]))
  if (name == "read.abares") {
    description <- "Provides simple downloading, parsing and importing of Australian
 Bureau of Agricultural and Resource Economics and Sciences (ABARES) data sources"
  }
  description <- cran_unquote(sub("\\.$", "", description))
  list(name = name, description = description)
}

get_user_text <- function(issue) {
  info <- gh::gh("GET /users/{username}", username = issue$user$login)
  name <- info$name %||% issue$user$login

  if (name == "Ryan Zomorrodi") {
    return("[Ryan Zomorrodi](https://ryanzomorrodi.github.io/)")
  }

  if (name == "Marc Burri") {
    return("[Marc Burri](https://marcburri.github.io/)")
  }

  url <- if (nzchar(info$blog)) info$blog else info$html_url
  if (!grepl("^https?:", url)) url <- paste0("http://", url)
  sprintf("[%s](%s)", name, url)
  
}

tidy_issue <- function(issue) {
  labels <- purrr::map_chr(issue$labels, "name")
  label <- labels[grepl("[0-9]\\/.*", labels)][1]
  df <- tibble::tibble(
    label = label,
    name = get_description(issue$body)$name,
    description = get_description(issue$body)$description,
    title = issue$title,
    holding = "holding" %in% purrr::map_chr(issue$labels, "name"),
    others = toString(purrr::map_chr(issue$labels, "name")),
    closed_at = issue$closed_at %||% NA,
    url = issue$html_url,
    user = get_user_text(issue),
    stats = dplyr::if_else("stats" %in% purrr::map_chr(issue$labels, "name"), " (Stats).", "")
  )
  
  dplyr::rowwise(df) %>%
    dplyr::mutate(text = sprintf("    * [%s](%s), %s. Submitted by %s. %s", name, url, description, user, stats))
}

get_issues <- function(label, state) {
  issues <- gh::gh(
    "GET /repos/{owner}/{repo}/issues",
    owner = "ropensci",
    repo = "software-review",
    state = state,
    .limit = Inf,
    .progress = FALSE
  )

  issues <- purrr::keep(issues, \(x) label %in% purrr::map_chr(x[["labels"]], "name"))
  
  purrr::map_df(issues, tidy_issue)
}
  
active_issues <- purrr::map_df(
  c("1/editor-checks","2/seeking-reviewer(s)","3/reviewer(s)-assigned","4/review(s)-in-awaiting-changes","5/awaiting-reviewer(s)-response","6/approved"),
  get_issues,
  state = "open"
)

closed_issues <- get_issues(state = "closed", label  ="6/approved")

ok_date <- function(date) {
  if (is.na(date)) {
    return(TRUE)
  } 
  
  anytime::anytime(date) >= last_newsletter
}

closed_issues <- dplyr::rowwise(closed_issues) %>%
  dplyr::filter(ok_date(closed_at))

issues <- dplyr::bind_rows(active_issues, closed_issues)


no_holding <- sum(issues$holding)
issues <- dplyr::filter(issues, !holding)
text <- sprintf("There are %s recently closed and active submissions", english(nrow(issues)))
if (no_holding > 0) {
  text <- paste0(
    text,
    sprintf(
      " and %s submission%s on hold.",
      no_holding,
      if (no_holding > 1) "s" else ""
    )
  )
} else {
  text <- paste0(text, ".")
}

count_label <- function(label) {
  no <- snakecase::to_sentence_case(english(sum(issues$label == label, na.rm = TRUE)))
  url <- glue::glue('https://github.com/ropensci/software-review/issues?q=is%3Aissue+is%3Aopen+sort%3Aupdated-desc+label%3A%22{label}%22')
  sprintf("* %s at ['%s'](%s):\n\n %s", no, label, url, paste0(issues$text[!is.na(issues$label)][ issues$label == label], collapse = "\n\n"))
}

cat(text)
cat(
  paste0(
    " Issues are at different stages: \n\n",
    paste0(
      purrr::map_chr(sort(unique(issues$label[!is.na(issues$label)]), decreasing = TRUE), count_label),
      collapse = "\n\n"
    )
  )
)
```

Find out more about [Software Peer Review](/software-review) and how to get involved.

## On the blog

<!-- Do not forget to rebase your branch! -->

```{r blog}

parse_one_post <- function(path){
  lines <- suppressWarnings(readLines(path, encoding = "UTF-8"))

  if (lines[2] == "headless: true") {
    return(NULL)
  }

  yaml <- blogdown:::split_yaml_body(lines)$yaml
  yaml <- glue::glue_collapse(yaml, sep = "\n")
  yaml <- yaml::yaml.load(yaml)
  
  language <- function(path) {
    name <- fs::path_ext_remove(fs::path_file(path))
    if (grepl("\\.[a-z][a-z]", name)) {
      sub(".*\\.", "", name)
    } else {
      "en"
    }
  }

  
  meta <- tibble::tibble(
    date = anytime::anydate(yaml$date),
    author = nice_string(yaml$author),
    title = yaml$title,
    software_peer_review = "Software Peer Review" %in% yaml$tags,
    tech_note = "tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    other = !"tech notes" %in% yaml$tags && !"Software Peer Review" %in% yaml$tags,
    socialImg = yaml$socialImg %||% "",
    socialAlt = yaml$socialAlt %||% "",
    description = yaml$description %||% "",
    newsletter = "newsletter" %in% yaml$tags,
    slug = yaml$slug,
    dir = fs::path_dir(path),
    language = language(path)
    )

  post_url <- if (meta[["language"]] == "en") {
    sprintf(
      "/blog/%s/%s/%s/%s",
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  } else {
    sprintf(
      "/%s/blog/%s/%s/%s/%s",
      meta[["language"]],
      lubridate::year(meta$date),
      stringr::str_pad(lubridate::month(meta$date), 2, "0", side = "left"),
      stringr::str_pad(lubridate::day(meta$date), 2, "0", side = "left"),
      meta$slug
    )
  }

  meta$url <- post_url
  meta
}
paths <- fs::dir_ls("..", recurse = TRUE, glob = "*.md")
paths <- paths[!paths %in% c("../_index.md",  "../2021-02-03-targets/raw_data_source.md",
  "../2021-02-03-targets/README.md")]
paths <- grep("index", paths, value = TRUE)
posts <- purrr::map_df(paths, parse_one_post)
posts <- dplyr::filter(posts, date >= as.Date(last_newsletter), !newsletter)
posts <- split(posts, posts[["dir"]])
format_post <- function(dir) {
  main_language <- if (any(dir[["language"]] == "en")) {
    "en"
  } else {
    dir[["language"]][[1]]
  }
  
  post <- dir[which(dir[["language"]] == main_language),]
  string <- sprintf("* [%s](%s) by %s", post$title, post$url, post$author)
  if (post$description != "") {
    string <- paste0(string, ". ", sub("[\\.\\!\\?]$", "", post$description), ".")
        if (url == "/blog/2025/09/18/markdown-programmatic-parsing/") {
      string <- paste(string, ".")
    }
  } else {
    string <- paste0(string, ".")  
  }
  
  if (post$socialImg != "") {
    if (!startsWith(post$socialImg, "https")) {
    img_file <- fs::path_file(post$socialImg)
    download.file(sprintf("https://ropensci.org/%s", post$socialImg), img_file)
    img_file %>% magick::image_read() %>% magick::image_scale("400x") %>% magick::image_write(img_file)
    } else {
      img_file <- post$socialImg
    }

    string <- paste0(
      string, "\n```{=html}\n",
      sprintf('{{< figure class="center" src="%s" alt="%s" width="400" >}}\n\n', img_file, post$socialAlt),
      "\n```\n"
    )
  }
  
other_langs <- dir[which(dir[["language"]] != main_language),]
  other_langs <- split(other_langs, sort(as.numeric(rownames(other_langs))))
  if (length(other_langs) > 0) {
    other_langs_text <- purrr::map_chr(
      other_langs,
      ~ sprintf("<a href='%s' lang='%s'>%s (%s)</a>", .x[["url"]], .x[["language"]], .x[["title"]], .x[["language"]])
      ) %>% 
      toString
    other_langs_text <- sprintf("Other languages: %s.", other_langs_text)
    string <- sprintf("%s %s", string, other_langs_text)
  }
  
  string
}
```

```{r, results='asis'}
software_review <- posts[purrr::map_lgl(posts, ~any(.x[["software_peer_review"]]))]
if (length(software_review) > 0) {
  cat("### Software Review\n\n")
  cat(
    paste0(
      purrr::map_chr(software_review, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}

others <- posts[purrr::map_lgl(posts, ~any(.x[["other"]]))]
if (length(others) > 0) {
  cat(
    paste0(
      purrr::map_chr(others, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}


tech_notes <- posts[purrr::map_lgl(posts, ~any(.x[["tech_note"]]))]
if (length(tech_notes) > 0) {
  cat("\n\n")
  cat("### Tech Notes\n\n")
  cat(
    paste0(
      purrr::map_chr(tech_notes, format_post),
      collapse = "\n\n"
    )
  )
  cat("\n\n")
}
```

## Calls for contributions

### Calls for maintainers

If you're interested in maintaining any of the R packages below, you might enjoy reading our blog post [What Does It Mean to Maintain a Package?](/blog/2023/02/07/what-does-it-mean-to-maintain-a-package/).


[photosearcher](https://docs.ropensci.org/photosearcher), searches Flickr for photographs and metadata. [Issue for volunteering](https://github.com/ropensci/photosearcher/issues/22).


### Calls for contributions

Refer to our [help wanted page](/help-wanted/) -- before opening a PR, we recommend asking in the issue whether help is still needed.


## Package development corner

Some useful tips for R package developers. :eyes:

### Update of the vcr chapter in HTTP testing in R

We've updated the introduction to vcr in the [HTTP testing in R book](https://books.ropensci.org/http-testing/vcr.html), so that it reflects the current vcr version, in particular the `local_cassette()` function.

### Jarl: just another R linter

Etienne Bacher introduced the new CLI jarl in a [blog post](https://www.etiennebacher.com/posts/2025-11-20-introducing-jarl/).
Like Air, it is run in the terminal.
Like the flir R package, it can fix bad patterns in R code.

### pkgdown 2.2.0 makes it easier for LLMs to read your documentation

The latest pkgdown version adds a step to the website building, that creates:

- a Markdown version of all HTML pages;
- a standard file called `llms.txt` that compiles the content of the homepage and of the reference and article indices.

Read more about this, including how to opt-out (one config line) in the [release announcement](https://tidyverse.org/blog/2025/11/pkgdown-2-2-0/).

### testthat 3.3.0

If you use testthat, don't miss its [release announcement](https://tidyverse.org/blog/2025/11/testthat-3-3-0/)! 
Of note:

- A new function `skip_unless_r()` to skip tests depending on the R version;
- Tests using Suggested packages will now be automatically skipped on CRAN;
- A new expectation `expect_shape()` to check the dimensions, number of rows or number of columns!

### A Git trick: keeping Git blame informative

If you reformat your whole codebase with Air, you'll end up with one or several commits that only change the aspect of your code.
A good code style is tantamount to readability, but when you're exploring the history through Git blame, that commit or those commits are not relevant.
This problem is solvable!

Thanks to Hugo Gruson to telling us this:
you can create (and .Rbuildignore) a file called `.git-blame-ignore-revs` where you list the hashes of the commits you want to exclude from Git blame.

Example in [testthat](https://github.com/r-lib/testthat/blob/d695714d301819cde6013ecd40a9116880f394ad/.git-blame-ignore-rev):

```txt
# This file lists revisions of large-scale formatting/style changes so that
# they can be excluded from git blame results.
#
# To set this file as the default ignore file for git blame, run:
#   $ git config blame.ignoreRevsFile .git-blame-ignore-revs

# https://github.com/r-lib/testthat/pull/2121
13d17788e5d3a54fa83beed25e325703608f8b9f
```

To use this file,

- you tell your local Git about it by running `git config blame.ignoreRevsFile .git-blame-ignore-revs`;
- GitHub, on the other hand, automatically [picks it up](https://docs.github.com/en/repositories/working-with-files/using-files/viewing-and-understanding-files#ignore-commits-in-the-blame-view)!

### Another Git trick: rebase all commits

Imagine you work on a private repository, on a single branch, not bothering about a clean Git history.
Then you have to make the repository public.
How to fix up the history before doing so?

You can rebase all the commits on main (or any branch) using the `--root` option of `git rebase`. 
So you can type `git rebase -i --root` and, say, combine all the commits into a very deceitful "First commit"!

You can practice [`git rebase -i`](https://docs.ropensci.org/saperlipopette/reference/exo_rebase_i.html) with the saperlipopette R package.

### Some talks from posit::conf(2025)

Besides talks by community members (see HQ section), you can watch recordings of talks relevant to software development:

- [Can I trust this Package](https://youtu.be/g6QU16jOt0g) by Collin S. Gillespie.
- [Extending the horizons of R with Rust](https://youtu.be/bAA_O_gNhfI) by Andrés Quintero.
- The keynote talk [The Psychology of Technologists](https://www.youtube.com/watch?v=B-bMX_l3nkk) by Cat Hicks.
- [The Curse of Documentation](https://youtu.be/ML8z8xkqIA0) by Michael Chow.
- [Air - A blazingly fast R code formatter](https://youtu.be/DJVSEOjvwb8) by Davis Vaughan and Lionel Henry.
- [Enemies to lovers: How non-programmers can make sparks fly when using testthat during package development](https://youtu.be/GEP61GjwTjE) by Libby McKenna.

## Last words

Thanks for reading! If you want to get involved with rOpenSci, check out our [Contributing Guide](https://contributing.ropensci.org) that can help direct you to the right place, whether you want to make code contributions, non-code contributions, or contribute in other ways like sharing use cases.
You can also support our work through [donations](/donate).

If you haven't subscribed to our newsletter yet, you can [do so via a form](/news/). Until it's time for our next newsletter, you can keep in touch with us via our [website](/) and [Mastodon account](https://hachyderm.io/@rOpenSci).